---
title: "Data Science for Non-Life Insurance"
author: "Alexiane Bölcs, Milla Campajola, Stefan Upeski"
date: "07 May 2026"
output: html_document
---

# Claims Development and Reserving

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(readxl)       
  library(dplyr)    
  library(tidyr)        
  library(purrr)   
  library(ggplot2)      
  library(scales)       
  library(randomForest) 
  library(survival)     
})

theme_set(theme_minimal(base_size = 11))
options(scipen = 999)

set.seed(42) 

# Directory for saved figures
dir.create("figs", showWarnings = FALSE)
```

## Data load and sanity check

```{r load, echo=TRUE}
path_data <- "main_data_groupD.xlsx"
claims <- read_excel(path_data) %>%
  mutate(
    Acc_Yr               = as.integer(Acc_Yr),
    Rpt_Yr_Lag           = as.integer(Rpt_Yr_Lag),
    claim_num            = as.character(claim_num),
    Dev_Yr               = as.integer(Dev_Yr),
    Cal_Yr               = as.integer(Cal_Yr),
    open_ind             = as.integer(open_ind),
    Accum_Infl_Index_Mid = as.numeric(Accum_Infl_Index_Mid),
    Incr_Payment         = as.numeric(Incr_Payment),
    Acc_Incr_Payment     = as.numeric(Acc_Incr_Payment)
  )

stopifnot(all(claims$Cal_Yr == claims$Acc_Yr + claims$Dev_Yr - 1))
stopifnot(!anyNA(claims))

cat(sprintf("Rows: %d | Claims: %d | AY: %d-%d | Dev_Yr: 1-%d\n",
            nrow(claims), n_distinct(claims$claim_num),
            min(claims$Acc_Yr), max(claims$Acc_Yr),
            max(claims$Dev_Yr)))

max_dev     <- max(claims$Dev_Yr)
max_cal     <- max(claims$Cal_Yr)
max_obs_cal <- max_cal                       # used in IBNR/RF sections
ay_range    <- sort(unique(claims$Acc_Yr))
```

The overall aim is to project reserves for the accident year 2009-2023.

# Task 2

### Exploratory Data Analysis

```{r EDA, echo=TRUE}
# 1. Reporting-delay distribution
p_delay <- claims %>%
  distinct(claim_num, Rpt_Yr_Lag) %>%
  count(Rpt_Yr_Lag) %>%
  mutate(pct = 100 * n / sum(n)) %>%
  ggplot(aes(factor(Rpt_Yr_Lag), pct)) +
  geom_col(fill = "darkorange") +
  geom_text(aes(label = sprintf("%.1f%%", pct)), vjust = -0.3, size = 3) +
  labs(title = "Reporting Delay Distribution",
       x = "Rpt_Yr_Lag", y = "% of claims")

ggsave("figs/reporting_delay.pdf", p_delay, width = 6, height = 4)
print(p_delay)

# 2. Incremental triangle heatmap
triangle_incr <- claims %>%
  group_by(Acc_Yr, Dev_Yr) %>%
  summarise(S = sum(Incr_Payment, na.rm = TRUE), .groups = "drop")

p_heat <- triangle_incr %>%
  ggplot(aes(factor(Dev_Yr), factor(Acc_Yr), fill = S / 1e6)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "C", labels = comma) +
  scale_y_discrete(limits = rev) +
  labs(title = "Incremental Claims Triangle (USD millions)",
       x = "Development Year", y = "Accident Year", fill = "M")

ggsave("figs/triangle_heatmap.pdf", p_heat, width = 7, height = 5)
print(p_heat)

# 3. Claim super-inflation
# Log-linear regression of median deflated payment on calendar year
# The slope (in log units per year) tells us the residual trend above
p_super <- claims %>%
  filter(Incr_Payment > 0) %>%
  mutate(defl = Incr_Payment / Accum_Infl_Index_Mid) %>%
  group_by(Cal_Yr) %>%
  summarise(mean_d = mean(defl), med_d = median(defl), .groups = "drop") %>%
  pivot_longer(-Cal_Yr, names_to = "stat", values_to = "val") %>%
  ggplot(aes(Cal_Yr, val, color = stat)) +
  geom_line(linewidth = 1) + geom_point() +
  scale_y_log10(labels = comma) +
  labs(title = "Deflated payments by calendar year",
       subtitle = "Upward slope = claim super-inflation above general index",
       y = "Deflated payment (log scale, USD)", x = "Calendar Year")

ggsave("figs/deflated_trend.pdf", p_super, width = 6, height = 4)
print(p_super)

# Formal estimate: log-linear regression on median deflated payment
super_fit <- claims %>%
  filter(Incr_Payment > 0) %>%
  mutate(defl = Incr_Payment / Accum_Infl_Index_Mid) %>%
  group_by(Cal_Yr) %>%
  summarise(med = median(defl), .groups = "drop") %>%
  mutate(log_med = log(med)) %>%
  { lm(log_med ~ Cal_Yr, data = .) }

# g_log is the log-rate (coefficient from lm); report it as-is
g_log <- coef(super_fit)[["Cal_Yr"]]
cat(sprintf("Claim super-inflation (log-linear on median): %+.2f%% /yr, R2=%.3f\n",
            100 * g_log, summary(super_fit)$r.squared))

# 4. Claim-size distribution (positive payments)
# We use log10 so the heavy upper tail is visible
pos_pay <- claims %>%
  filter(Incr_Payment > 0) %>%
  mutate(log10_pay = log10(Incr_Payment))

q_med <- median(pos_pay$Incr_Payment)
q_99  <- quantile(pos_pay$Incr_Payment, 0.99)
share_below_1k <- mean(pos_pay$Incr_Payment < 1000)

cat(sprintf("\nPositive payments: median = %s USD, 99th pct = %s USD\n",
            format(round(q_med), big.mark = ","),
            format(round(q_99), big.mark = ",")))
cat(sprintf("Share of positive payments below 1,000 USD: %.1f%%\n",
            100 * share_below_1k))
cat(sprintf("Share of TOTAL dollars from top 1%% of payments: %.1f%%\n",
            100 * sum(pos_pay$Incr_Payment[pos_pay$Incr_Payment >= q_99]) /
              sum(pos_pay$Incr_Payment)))
p_size <- pos_pay %>%
  ggplot(aes(log10_pay)) +
  geom_histogram(aes(y = after_stat(density)), bins = 60,
                 fill = "#3498db", color = "white", alpha = 0.85) +
  geom_density(color = "#1b4f72", linewidth = 0.8) +
  geom_vline(xintercept = log10(q_med), linetype = "dashed",
             color = "#c0392b", linewidth = 0.8) +
  geom_vline(xintercept = log10(q_99),  linetype = "dashed",
             color = "#27ae60", linewidth = 0.8) +
  annotate("text", x = log10(q_med), y = Inf, vjust = 1.3, hjust = -0.05,
           label = sprintf("median\n%s", format(round(q_med), big.mark = ",")),
           color = "#c0392b", size = 3) +
  annotate("text", x = log10(q_99),  y = Inf, vjust = 1.3, hjust = -0.05,
           label = sprintf("99th pct\n%s", format(round(q_99), big.mark = ",")),
           color = "#27ae60", size = 3) +
  scale_x_continuous(breaks = 0:7,
                     labels = c("$1", "$10", "$100", "$1k", "$10k",
                                "$100k", "$1M", "$10M")) +
  labs(title = "Distribution of positive incremental payments (log10 scale)",
       subtitle = "Long right tail: top 1% of payments drive a large fraction of total volume",
       x = "Payment amount (log10 USD)", y = "Density")
ggsave("figs/claim_size.pdf", p_size, width = 7, height = 4)
print(p_size)

# 5. Severity by Accident Year (nominal vs deflated)
# Mean positive payment per claim by accident year, both NOMINAL and DEFLATED
# We aggregate by accident year (not by calendar year) so each bar
# represents the average severity of claims OCCURRED in that year
sev_by_ay <- claims %>%
  filter(Incr_Payment > 0) %>%
  mutate(defl = Incr_Payment / Accum_Infl_Index_Mid) %>%
  group_by(Acc_Yr) %>%
  summarise(
    n_pay   = n(),
    mean_nom = mean(Incr_Payment),
    mean_def = mean(defl),
    .groups = "drop"
  )
cat("\nMean positive payment per accident year:\n")
print(sev_by_ay %>% mutate(mean_nom = round(mean_nom),
                           mean_def = round(mean_def)))
p_sev_ay <- sev_by_ay %>%
  select(Acc_Yr,
         "Nominal (USD)"            = mean_nom,
         "Deflated (USD, real)"     = mean_def) %>%
  pivot_longer(-Acc_Yr, names_to = "Type", values_to = "Severity") %>%
  ggplot(aes(factor(Acc_Yr), Severity, fill = Type)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("Nominal (USD)"        = "#1b4f72",
                               "Deflated (USD, real)" = "#e67e22")) +
  scale_y_continuous(labels = comma) +
  labs(title = "Mean positive payment by accident year — nominal vs deflated",
       subtitle = "Gap between bars = general inflation; 
       trend in orange = real-terms claim severity",
       x = "Accident Year", y = "Mean payment (USD)") +
  theme(legend.position = "bottom")
ggsave("figs/fig5_severity_by_ay.pdf", p_sev_ay, width = 7, height = 4)
print(p_sev_ay)

# 6. Closure dynamics by accident year
# Share of claims still open at the end of each development year, by
# accident year cohort. We restrict to mature cohorts (Acc_Yr <= 2015)
# so that the curves can be observed up to Dev_Yr ~ 8-15 without
# right-censoring biasing the picture
mature_ay <- 2009:2015
closure_curves <- claims %>%
  filter(Acc_Yr %in% mature_ay) %>%
  group_by(Acc_Yr, Dev_Yr) %>%
  summarise(open_share = mean(open_ind == 1), .groups = "drop")

p_closure <- closure_curves %>%
  ggplot(aes(Dev_Yr, 100 * open_share, color = factor(Acc_Yr))) +
  geom_line(linewidth = 0.8) + geom_point(size = 1.5) +
  scale_x_continuous(breaks = 1:max_dev) +
  scale_color_viridis_d(option = "D", end = 0.9) +
  labs(title = "Share of claims still OPEN by development year (mature cohorts)",
       subtitle = "AYs 2009-2015: motivates the survival/Cox correction in Section 10",
       x = "Development Year",
       y = "% of claims still open",
       color = "Accident Year") +
  theme(legend.position = "right")
ggsave("figs/fig6_closure_curves.pdf", p_closure, width = 7, height = 4)
print(p_closure)
```

# Task 1

## Chain Ladder Unadjusted

```{r CL unadjusted, echo=TRUE}
# Build full Acc_Yr × Dev_Yr rectangle
# obs = TRUE for cells we actually observe (Cal_Yr <= max_cal = 2023)
triangle_full <- expand_grid(Acc_Yr = ay_range, Dev_Yr = 1:max_dev) %>%
  mutate(Cal_Yr = Acc_Yr + Dev_Yr - 1,
         obs    = Cal_Yr <= max_cal) %>%
  left_join(triangle_incr, by = c("Acc_Yr", "Dev_Yr")) %>%
  mutate(S = ifelse(obs, replace_na(S, 0), NA_real_))

# Cumulative triangle (within observed cells only)
triangle_cum <- triangle_full %>%
  group_by(Acc_Yr) %>% arrange(Dev_Yr, .by_group = TRUE) %>%
  mutate(C = ifelse(obs, cumsum(replace_na(S, 0)), NA_real_)) %>%
  ungroup()

# Volume-weighted (Mack) development factors
dev_factors <- map_dfr(1:(max_dev - 1), function(k) {
  tmp <- triangle_cum %>%
    filter(Dev_Yr %in% c(k, k + 1), obs) %>%
    select(Acc_Yr, Dev_Yr, C) %>%
    pivot_wider(names_from = Dev_Yr, values_from = C, names_prefix = "d") %>%
    filter(!is.na(.data[[paste0("d", k)]]),
           !is.na(.data[[paste0("d", k + 1)]]),
           .data[[paste0("d", k)]] > 0)
  tibble(k = k,
         f_k = sum(tmp[[paste0("d", k + 1)]]) / sum(tmp[[paste0("d", k)]]))
})
tail_factor_nominal_raw <- dev_factors$f_k[dev_factors$k == max_dev - 1]
cat(sprintf("Raw last development factor before truncation: %.6f\n",
            tail_factor_nominal_raw))

# Tail factor f_14 = 1: claims are essentially closed at Dev 14
dev_factors$f_k[dev_factors$k == max_dev - 1] <- 1

# Cumulative factors F_k (dev k to ultimate)
F_k <- tibble(Dev_Yr = 1:max_dev) %>%
  mutate(F_k = sapply(Dev_Yr, function(k)
    if (k == max_dev) 1 else prod(dev_factors$f_k[dev_factors$k >= k])))

# Reserve per accident year
latest_obs <- triangle_cum %>%
  filter(obs, !is.na(C)) %>%
  group_by(Acc_Yr) %>%
  summarise(k_obs = max(Dev_Yr),
            C_obs = C[Dev_Yr == max(Dev_Yr)],
            .groups = "drop")
reserve_unadj <- latest_obs %>%
  left_join(F_k, by = c("k_obs" = "Dev_Yr")) %>%
  mutate(C_ult = C_obs * F_k,
         R_hat = C_ult - C_obs)
cat("\n=== CL unadjusted reserves ===\n")
print(reserve_unadj)
cat(sprintf("Total unadjusted IBNR: %s USD\n",
            format(round(sum(reserve_unadj$R_hat)), big.mark = ",")))
```

### Verifying assumption of Chain Ladder (Mack model)

```{r assumption CL, echo=TRUE}
# We complement 3 Chain-Ladder model assumptions:
# M1: E[C_{i,k+1} | C_{i,k}] = f_k * C_{i,k}            
# M2: Var[C_{i,k+1} | C_{i,k}] = sigma_k^2 * C_{i,k}    
# M3: accident years are independent
#
# We compute:
# Mack's sigma_k^2 estimator
# Properly standardised residuals z_{i,k}
# z_{i,k} = (C_{i,k+1} - f_k * C_{i,k}) / (sigma_k * sqrt(C_{i,k}))
# which under M1+M2 should be ~N(0,1) with no pattern
# Mack's rank-correlation test for accident-year independence

# Build link_ratios and residuals_df from triangle_cum
# Pairs (C_{i,k}, C_{i,k+1}) that are both observed
link_ratios <- triangle_cum %>%
  filter(obs, !is.na(C)) %>%
  select(Acc_Yr, Dev_Yr, C) %>%
  arrange(Acc_Yr, Dev_Yr) %>%
  group_by(Acc_Yr) %>%
  mutate(C_k1 = lead(C)) %>%
  ungroup() %>%
  filter(!is.na(C_k1), Dev_Yr < max_dev) %>%
  rename(C_k = C) %>%
  mutate(link_ratio = C_k1 / C_k) %>%
  left_join(dev_factors %>% rename(Dev_Yr = k), by = "Dev_Yr") %>%
  filter(!is.na(f_k))

residuals_df <- link_ratios %>%
  mutate(fitted   = f_k * C_k,
         residual = C_k1 - fitted)

# Mack sigma_k^2
# sigma_k^2 = 1/(I-k-1) * sum_i C_{i,k} * (link_ratio_{i,k} - f_k)^2
sigma2_mack <- map_dfr(sort(unique(link_ratios$Dev_Yr)), function(k) {
  sub   <- link_ratios %>% filter(Dev_Yr == k)
  f_hat <- dev_factors$f_k[dev_factors$k == k]
  I_k   <- nrow(sub)
  if (I_k <= 1) return(tibble(Dev_Yr = k, sigma2_k = NA_real_, n_pairs = I_k))
  s2 <- sum(sub$C_k * (sub$link_ratio - f_hat)^2) / (I_k - 1)
  tibble(Dev_Yr = k, sigma2_k = s2, n_pairs = I_k)
})

cat("\n=== Mack sigma_k^2 estimates ===\n")
print(sigma2_mack %>% mutate(sigma_k = sqrt(sigma2_k)))

# Properly standardised Mack residuals
mack_resid <- residuals_df %>%
  left_join(sigma2_mack %>% select(Dev_Yr, sigma2_k), by = "Dev_Yr") %>%
  mutate(z_mack = (C_k1 - fitted) / (sqrt(sigma2_k) * sqrt(C_k)))

# Plot Mack-standardised residuals vs fitted
p_mack_resid <- mack_resid %>%
  filter(!is.na(z_mack)) %>%
  ggplot(aes(fitted / 1e6, z_mack)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = c(-2, 0, 2),
             linetype = c("dashed", "solid", "dashed"),
             color    = c("grey50", "black", "grey50")) +
  labs(title    = "Mack-standardised residuals z_{i,k} vs fitted",
       subtitle = "Should look like ~N(0,1): no funnel, no trend, ~95% inside +/-2",
       x = "Fitted cumulative payment (USD millions)",
       y = "Mack standardised residual")

if (!dir.exists("figs")) dir.create("figs", recursive = TRUE)
ggsave("figs/mack_residuals.pdf", p_mack_resid, width = 7, height = 4)
print(p_mack_resid)

# Sanity numbers: mean and sd of Mack residuals should be ~0 and ~1
cat("\n=== Mack residual sanity check ===\n")
cat(sprintf("  mean(z_mack) = %.3f   (target 0)\n",
            mean(mack_resid$z_mack, na.rm = TRUE)))
cat(sprintf("  sd  (z_mack) = %.3f   (target 1)\n",
            sd(mack_resid$z_mack, na.rm = TRUE)))
cat(sprintf("  %% within +/-2: %.1f%%   (target ~95%%)\n",
            100 * mean(abs(mack_resid$z_mack) <= 2, na.rm = TRUE)))

# Mack's rank-correlation test for accident-year independence
# For each k with >= 4 obs, Spearman(C_{i,k}, link_ratio_{i,k})
mack_indep <- map_dfr(sort(unique(link_ratios$Dev_Yr)), function(k) {
  sub <- link_ratios %>% filter(Dev_Yr == k)
  if (nrow(sub) < 4) return(tibble(Dev_Yr = k, rho = NA_real_,
                                   p_val = NA_real_, n = nrow(sub)))
  test <- suppressWarnings(cor.test(sub$C_k, sub$link_ratio, method = "spearman"))
  tibble(Dev_Yr = k, rho = unname(test$estimate),
         p_val = test$p.value, n = nrow(sub))
})

cat("\n=== Mack rank-correlation test (M3 — AY independence) ===\n")
cat("  Spearman rho between C_{i,k} and link_ratio_{i,k}\n")
cat("  Many small p-values would suggest violation of independence\n")
print(mack_indep %>% mutate(across(c(rho, p_val), ~ round(., 3))))
```

## Chain Ladder Inflation-Adjusted

```{r CL inflation-adjusted, echo=TRUE}
# Inflation index table (one row per calendar year)
infl_tbl <- claims %>% distinct(Cal_Yr, Accum_Infl_Index_Mid) %>% arrange(Cal_Yr)

# Deflated cumulative triangle
tri_defl <- triangle_full %>%
  left_join(infl_tbl, by = "Cal_Yr") %>%
  mutate(S_def = ifelse(obs, S / Accum_Infl_Index_Mid, NA_real_)) %>%
  group_by(Acc_Yr) %>% arrange(Dev_Yr, .by_group = TRUE) %>%
  mutate(C_def = ifelse(obs, cumsum(replace_na(S_def, 0)), NA_real_)) %>%
  ungroup()

# Development factors on the deflated triangle
f_def <- map_dfr(1:(max_dev - 1), function(k) {
  tmp <- tri_defl %>%
    filter(Dev_Yr %in% c(k, k + 1), obs) %>%
    select(Acc_Yr, Dev_Yr, C_def) %>%
    pivot_wider(names_from = Dev_Yr, values_from = C_def, names_prefix = "d") %>%
    filter(!is.na(.data[[paste0("d", k)]]),
           !is.na(.data[[paste0("d", k + 1)]]),
           .data[[paste0("d", k)]] > 0)
  tibble(k = k,
         f_def = sum(tmp[[paste0("d", k + 1)]]) / sum(tmp[[paste0("d", k)]]))
})
f_def$f_def[f_def$k == max_dev - 1] <- 1

# Project deflated rectangle one step at a time
rect_def <- tri_defl %>% select(Acc_Yr, Dev_Yr, Cal_Yr, obs, C_def)
for (ay in ay_range) for (k in 1:(max_dev - 1)) {
  ic  <- which(rect_def$Acc_Yr == ay & rect_def$Dev_Yr == k)
  in_ <- which(rect_def$Acc_Yr == ay & rect_def$Dev_Yr == (k + 1))
  if (!is.na(rect_def$C_def[ic]) && is.na(rect_def$C_def[in_]))
    rect_def$C_def[in_] <- rect_def$C_def[ic] * f_def$f_def[f_def$k == k]
}

# Extrapolate inflation index (3-year trailing mean rate) 
last_rates <- diff(infl_tbl$Accum_Infl_Index_Mid) /
  head(infl_tbl$Accum_Infl_Index_Mid, -1)
avg_rate   <- mean(tail(last_rates, 3))
cat(sprintf("Extrapolating inflation at %.2f%%/yr (3-year trailing mean)\n",
            100 * avg_rate))

max_cal_future <- max(ay_range) + max_dev - 1
infl_full <- tibble(Cal_Yr = min(infl_tbl$Cal_Yr):max_cal_future) %>%
  left_join(infl_tbl, by = "Cal_Yr")
for (i in 2:nrow(infl_full))
  if (is.na(infl_full$Accum_Infl_Index_Mid[i]))
    infl_full$Accum_Infl_Index_Mid[i] <-
  infl_full$Accum_Infl_Index_Mid[i - 1] * (1 + avg_rate)

# Re-inflate and compute reserve
reserve_infl <- rect_def %>%
  group_by(Acc_Yr) %>% arrange(Dev_Yr, .by_group = TRUE) %>%
  mutate(S_def_proj = C_def - lag(C_def, default = 0)) %>%
  ungroup() %>%
  filter(!obs) %>%
  left_join(infl_full, by = "Cal_Yr") %>%
  mutate(S_nom_proj = S_def_proj * Accum_Infl_Index_Mid) %>%
  group_by(Acc_Yr) %>%
  summarise(R_hat_infl = sum(S_nom_proj, na.rm = TRUE), .groups = "drop")

cat("\n=== CL inflation-adjusted reserves ===\n")
print(reserve_infl)
cat(sprintf("Total inflation-adjusted CL Reserve: %s USD\n",
            format(round(sum(reserve_infl$R_hat_infl)), big.mark = ",")))

# Development factors comparison: nominal vs deflated
p_factors <- dev_factors %>%
  rename(f_nominal = f_k) %>%
  left_join(f_def %>% rename(f_deflated = f_def), by = "k") %>%
  pivot_longer(-k, names_to = "Type", values_to = "Factor") %>%
  ggplot(aes(k, Factor, color = Type, group = Type)) +
  geom_line(linewidth = 1) + geom_point(size = 2) +
  scale_color_manual(values = c(f_nominal  = "#1b4f72",
                                f_deflated = "#c0392b"),
                     labels = c(f_nominal  = "Nominal (unadjusted)",
                                f_deflated = "Deflated (inflation-adj.)")) +
  scale_x_continuous(breaks = 1:(max_dev - 1)) +
  labs(title = "Development factors: nominal vs inflation-adjusted",
       subtitle = "Deflated factors are uniformly smaller — confirms inflation in the data",
       x = "Development step k", y = "f_k") +
  theme(legend.position = "bottom")
ggsave("figs/dev_factors.pdf", p_factors, width = 7, height = 4)
print(p_factors)
```

### Verifying assumption of Chain Ladder (Mack model) on inflation-adjusted triangle

```{r assumption CL inflation-adjusted, echo=TRUE}
# Build link_ratios_def and residuals_def from tri_defl
link_ratios_def <- tri_defl %>%
  filter(obs, !is.na(C_def)) %>%
  select(Acc_Yr, Dev_Yr, C_def) %>%
  arrange(Acc_Yr, Dev_Yr) %>%
  group_by(Acc_Yr) %>%
  mutate(C_def_k1 = lead(C_def)) %>%
  ungroup() %>%
  filter(!is.na(C_def_k1), Dev_Yr < max_dev) %>%
  rename(C_def_k = C_def) %>%
  mutate(link_ratio = C_def_k1 / C_def_k) %>%
  left_join(f_def %>% rename(Dev_Yr = k), by = "Dev_Yr") %>%
  filter(!is.na(f_def))
residuals_def <- link_ratios_def %>%
  mutate(fitted   = f_def * C_def_k,
         residual = C_def_k1 - fitted)

# Mack sigma_k^2 on deflated triangle
sigma2_mack_def <- map_dfr(sort(unique(link_ratios_def$Dev_Yr)), function(k) {
  sub   <- link_ratios_def %>% filter(Dev_Yr == k)
  f_hat <- f_def$f_def[f_def$k == k]
  I_k   <- nrow(sub)
  if (I_k <= 1) return(tibble(Dev_Yr = k, sigma2_k = NA_real_, n_pairs = I_k))
  s2 <- sum(sub$C_def_k * (sub$link_ratio - f_hat)^2) / (I_k - 1)
  tibble(Dev_Yr = k, sigma2_k = s2, n_pairs = I_k)
})
cat("\n=== Mack sigma_k^2 estimates (deflated) ===\n")
print(sigma2_mack_def %>% mutate(sigma_k = sqrt(sigma2_k)))

# Properly standardised Mack residuals on deflated triangle
mack_resid_def <- residuals_def %>%
  left_join(sigma2_mack_def %>% select(Dev_Yr, sigma2_k), by = "Dev_Yr") %>%
  mutate(z_mack = (C_def_k1 - fitted) / (sqrt(sigma2_k) * sqrt(C_def_k)))
p_mack_resid_def <- mack_resid_def %>%
  filter(!is.na(z_mack)) %>%
  ggplot(aes(fitted / 1e6, z_mack)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = c(-2, 0, 2),
             linetype = c("dashed", "solid", "dashed"),
             color    = c("grey50", "black", "grey50")) +
  labs(title    = "Mack-standardised residuals — inflation-adjusted CL",
       subtitle = "Should look like ~N(0,1): no funnel, no trend, ~95% inside +/-2",
       x = "Fitted deflated cumulative payment (USD millions, base year)",
       y = "Mack standardised residual")
if (!dir.exists("figs")) dir.create("figs", recursive = TRUE)
ggsave("figs/mack_residuals_deflated.pdf",
       p_mack_resid_def, width = 7, height = 4)
print(p_mack_resid_def)
cat("\n=== Mack residual sanity check (deflated) ===\n")
cat(sprintf("  mean(z_mack) = %.3f   (target 0)\n",
            mean(mack_resid_def$z_mack, na.rm = TRUE)))
cat(sprintf("  sd  (z_mack) = %.3f   (target 1)\n",
            sd(mack_resid_def$z_mack, na.rm = TRUE)))
cat(sprintf("  %% within +/-2: %.1f%%   (target ~95%%)\n",
            100 * mean(abs(mack_resid_def$z_mack) <= 2, na.rm = TRUE)))
# Mack rank-correlation test on deflated triangle
mack_indep_def <- map_dfr(sort(unique(link_ratios_def$Dev_Yr)), function(k) {
  sub <- link_ratios_def %>% filter(Dev_Yr == k)
  if (nrow(sub) < 4) return(tibble(Dev_Yr = k, rho = NA_real_,
                                   p_val = NA_real_, n = nrow(sub)))
  test <- suppressWarnings(cor.test(sub$C_def_k, sub$link_ratio, method = "spearman"))
  tibble(Dev_Yr = k, rho = unname(test$estimate),
         p_val = test$p.value, n = nrow(sub))
})
cat("\n=== Mack rank-correlation test (deflated, M3 — AY independence) ===\n")
cat("  Spearman rho between C_def_{i,k} and link_ratio_def_{i,k}\n")
print(mack_indep_def %>% mutate(across(c(rho, p_val), ~ round(., 3))))

# Comparison: nominal vs deflated
if (exists("mack_resid") && exists("mack_indep")) {
  cat("\n=== Nominal vs deflated — summary comparison ===\n")
  cmp <- tibble(
    triangle    = c("nominal", "deflated"),
    mean_z      = c(mean(mack_resid$z_mack,     na.rm = TRUE),
                    mean(mack_resid_def$z_mack, na.rm = TRUE)),
    sd_z        = c(sd(mack_resid$z_mack,       na.rm = TRUE),
                    sd(mack_resid_def$z_mack,   na.rm = TRUE)),
    pct_in_2    = c(100 * mean(abs(mack_resid$z_mack)     <= 2, na.rm = TRUE),
                    100 * mean(abs(mack_resid_def$z_mack) <= 2, na.rm = TRUE)),
    n_signif_M3 = c(sum(mack_indep$p_val     < 0.05, na.rm = TRUE),
                    sum(mack_indep_def$p_val < 0.05, na.rm = TRUE))
  )
  print(cmp %>% mutate(across(where(is.numeric), ~ round(., 3))))
}
```

## Chain Ladder Back Test Validation

Done jointly with the Random Forest backtest. See further down.

## Chain Ladder Super-Inflation Sensitivity

We layer a super-inflation shock on top of the baseline projection, first at a 
moderate level and then at full strength, to gauge how the reserve estimate 
responds as the inflationary stress intensifies.

```{r CL super-inflation, echo=TRUE}
g_discrete  <- exp(g_log) - 1          
g_moderate  <- g_discrete / 2          # half-rate scenario
cat(sprintf("\nClaim super-inflation rates:\n"))
cat(sprintf("  Log-rate (lm coeff):  %+.4f /yr\n", g_log))
cat(sprintf("  Discrete (exp-1):     %+.2f%% /yr  <- used in multiplier\n",
            100 * g_discrete))
cat(sprintf("  Moderate (half):      %+.2f%% /yr\n", 100 * g_moderate))
max_cal <- max(claims$Cal_Yr)   # 2023

# Future incremental cells (deflated), from the inflation-adjusted CL
future_cells <- rect_def %>%
  arrange(Acc_Yr, Dev_Yr) %>%
  group_by(Acc_Yr) %>%
  mutate(S_def_proj = C_def - lag(C_def, default = 0)) %>%
  ungroup() %>%
  filter(!obs, S_def_proj > 0) %>%
  select(Acc_Yr, Dev_Yr, Cal_Yr, S_def_proj) %>%
  left_join(infl_full %>% select(Cal_Yr, Accum_Infl_Index_Mid),
            by = "Cal_Yr") %>%
  mutate(
    S_nom_base  = S_def_proj * Accum_Infl_Index_Mid,
    years_ahead = Cal_Yr - max_cal,
    M_moderate = (1 + g_moderate)^years_ahead,
    M_full     = (1 + g_discrete)^years_ahead,
    S_nom_mod  = S_nom_base * M_moderate,
    S_nom_full = S_nom_base * M_full
  )
reserves_scenarios <- future_cells %>%
  group_by(Acc_Yr) %>%
  summarise(R_base     = sum(S_nom_base, na.rm = TRUE),
            R_moderate = sum(S_nom_mod,  na.rm = TRUE),
            R_full     = sum(S_nom_full, na.rm = TRUE),
            .groups    = "drop")

# Sanity check: base must match reserve_infl
diff_base <- abs(sum(reserves_scenarios$R_base) -
                   sum(reserve_infl$R_hat_infl))
stopifnot(diff_base < 1)  
cat(sprintf("Sanity check — base vs reserve_infl: diff = $%.0f (OK)\n",
            diff_base))
total_base     <- sum(reserves_scenarios$R_base)
total_moderate <- sum(reserves_scenarios$R_moderate)
total_full     <- sum(reserves_scenarios$R_full)
cat(sprintf("\n=== Super-inflation scenarios ===\n"))
cat(sprintf("Scenario 0 — base (g = 0%%, trend stops):      %s USD\n",
            format(round(total_base),     big.mark = ",")))
cat(sprintf("Scenario 1 — moderate (g = %+.2f%%/yr):         %s USD\n",
            100 * g_moderate,
            format(round(total_moderate), big.mark = ",")))
cat(sprintf("Scenario 2 — full (g = %+.2f%%/yr):              %s USD\n",
            100 * g_discrete,
            format(round(total_full),     big.mark = ",")))
cat(sprintf("Sensitivity range: [%s; %s] USD\n",
            format(round(total_base), big.mark = ","),
            format(round(total_full), big.mark = ",")))
p_scenarios <- reserves_scenarios %>%
  select(Acc_Yr,
         "Scenario 0 — base (g = 0%)"            = R_base,
         "Scenario 1 — moderate (g/2)"            = R_moderate,
         "Scenario 2 — full super-inflation (g)"  = R_full) %>%
  pivot_longer(-Acc_Yr, names_to = "Scenario", values_to = "Reserve") %>%
  ggplot(aes(factor(Acc_Yr), Reserve / 1e6, fill = Scenario)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("Scenario 0 — base (g = 0%)"           = "#1b4f72",
                               "Scenario 1 — moderate (g/2)"           = "#e67e22",
                               "Scenario 2 — full super-inflation (g)" = "#c0392b")) +
  scale_y_continuous(labels = comma) +
  labs(title = "CL reserve sensitivity to claim super-inflation",
       x = "Accident Year", y = "Reserve (USD millions)") +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 8))
ggsave("figs/super_inflation.pdf", p_scenarios, width = 7, height = 4)
print(p_scenarios)
```

# Task 3

## 3-part Random Forest - RBNS projection

```{r 3 RF for RBNS, echo=TRUE}
# For each (claim, Dev_Yr) row we lag back one step to get the
# state at the start of the year. lag_open == 1 means the claim
# was open at the start, this is the condition we train on
claims_rf <- claims %>%
  arrange(claim_num, Dev_Yr) %>%
  group_by(claim_num) %>%
  mutate(
    lag_cum_payment  = lag(Acc_Incr_Payment, default = 0),
    lag_open         = lag(open_ind,         default = 1L),
    has_paid_before  = as.integer(lag_cum_payment > 0),
    log_lag_cum      = log1p(lag_cum_payment),
    dev_ratio        = Dev_Yr / max_dev,
    has_payment      = as.integer(Incr_Payment > 0),
    payment_defl     = Incr_Payment / Accum_Infl_Index_Mid,
    log_payment_defl = ifelse(payment_defl > 0, log(payment_defl), NA_real_),
    closed_this_yr   = as.integer(lag_open == 1L & open_ind == 0L)
  ) %>% ungroup()
feats <- c("Acc_Yr", "Dev_Yr", "Rpt_Yr_Lag",
           "has_paid_before", "log_lag_cum", "dev_ratio")
train <- claims_rf %>% filter(lag_open == 1)
# Fit the three random forests
rf_freq <- randomForest(
  x        = train[, feats],
  y        = as.factor(train$has_payment),
  ntree    = 200, mtry = 3, maxnodes = 512,
  sampsize = c("0" = 10000, "1" = 10000)
)
sev_tr  <- train %>% filter(has_payment == 1, !is.na(log_payment_defl))
rf_sev  <- randomForest(
  x    = sev_tr[, feats],
  y    = sev_tr$log_payment_defl,
  ntree = 200, mtry = 3, maxnodes = 512
)
# Residual variance for the smearing (log-normal retransformation) correction
sigma2 <- mean((rf_sev$y - rf_sev$predicted)^2, na.rm = TRUE)
rf_close <- randomForest(
  x        = train[, feats],
  y        = as.factor(train$closed_this_yr),
  ntree    = 200, mtry = 3, maxnodes = 512,
  sampsize = c("0" = 10000, "1" = 10000)
)
cat(sprintf("\nRF smearing correction — sigma^2 = %.4f\n", sigma2))
cat("RF feature importances:\n")
cat("  Frequency: "); print(round(importance(rf_freq) [, 1], 2))
cat("  Severity : "); print(round(importance(rf_sev)  [, 1], 2))
cat("  Closure  : "); print(round(importance(rf_close)[, 1], 2))

# OOB performance of the three RFs
# Out-of-Bag is randomForest's internal cross-validation:
# each tree is trained on a bootstrap sample, then predictions are
# made only on the rows not in that bootstrap. This gives an error estimate
cat("\n=== RF Out-of-Bag (OOB) performance ===\n")
# Frequency: OOB confusion matrix and error rate
freq_oob <- rf_freq$confusion
cat("Frequency RF — OOB confusion matrix:\n")
print(freq_oob)
cat(sprintf("  Overall OOB error rate: %.2f%%\n",
            100 * (1 - sum(diag(freq_oob[, 1:2])) / sum(freq_oob[, 1:2]))))
# Severity: % variance explained by OOB predictions
sev_var_expl <- 100 * (1 - mean((rf_sev$y - rf_sev$predicted)^2, na.rm = TRUE) /
                         var(rf_sev$y, na.rm = TRUE))
cat(sprintf("Severity RF  — %% variance explained (OOB): %.2f%%\n", sev_var_expl))
cat(sprintf("Severity RF  — OOB RMSE on log scale:      %.4f\n",
            sqrt(mean((rf_sev$y - rf_sev$predicted)^2, na.rm = TRUE))))
# Closure: OOB confusion matrix
close_oob <- rf_close$confusion
cat("Closure RF   — OOB confusion matrix:\n")
print(close_oob)
cat(sprintf("  Overall OOB error rate: %.2f%%\n",
            100 * (1 - sum(diag(close_oob[, 1:2])) / sum(close_oob[, 1:2]))))

# RBNS projection loop
# We carry a "survival probability" prob_open = P(still open at step k),
# updated multiplicatively after each development year
state <- claims_rf %>%
  group_by(claim_num) %>%
  filter(Dev_Yr == max(Dev_Yr)) %>% ungroup() %>%
  filter(open_ind == 1, Dev_Yr < max_dev) %>%
  transmute(claim_num, Acc_Yr, Rpt_Yr_Lag,
            dev       = Dev_Yr,
            cum       = Acc_Incr_Payment,
            prob_open = 1)
reserves_ml <- tibble(Acc_Yr = ay_range, R_hat_RF = 0)
for (step in 1:max_dev) {
  state <- state %>% filter(dev < max_dev)
  if (nrow(state) == 0) break
  d_next  <- state$dev + 1
  cy_next <- state$Acc_Yr + d_next - 1
  gen_idx <- infl_full$Accum_Infl_Index_Mid[match(cy_next, infl_full$Cal_Yr)]
  gen_idx[is.na(gen_idx)] <- tail(infl_full$Accum_Infl_Index_Mid, 1)
  X <- data.frame(
    Acc_Yr          = state$Acc_Yr,
    Dev_Yr          = d_next,
    Rpt_Yr_Lag      = state$Rpt_Yr_Lag,
    has_paid_before = as.integer(state$cum > 0),
    log_lag_cum     = log1p(state$cum),
    dev_ratio       = d_next / max_dev
  )[, feats]
  p_pay   <- predict(rf_freq,  X, type = "prob")[, "1"]
  log_sev <- predict(rf_sev,   X)
  p_close <- predict(rf_close, X, type = "prob")[, "1"]
  sev_nom <- exp(log_sev + sigma2 / 2) * gen_idx
  expected_nominal <- state$prob_open * p_pay * sev_nom
  tmp <- tibble(Acc_Yr = state$Acc_Yr, pay = expected_nominal) %>%
    group_by(Acc_Yr) %>% summarise(pay = sum(pay), .groups = "drop")
  reserves_ml <- reserves_ml %>%
    left_join(tmp, by = "Acc_Yr") %>%
    mutate(R_hat_RF = R_hat_RF + replace_na(pay, 0)) %>%
    select(-pay)
  state$cum       <- state$cum + expected_nominal
  state$prob_open <- state$prob_open * (1 - p_close)
  state$dev       <- d_next
}
cat("\n=== RF RBNS reserves ===\n")
print(reserves_ml)
cat(sprintf("Total RF RBNS reserve: %s USD\n",
            format(round(sum(reserves_ml$R_hat_RF)), big.mark = ",")))
```

### Pure IBNR via reporting-lag Distribution

```{r pure IBNR estimation, echo=TRUE}
# compute the lag distribution only on fully mature AYs (<= 2013)
# where every possible lag has had time to be observed
lag_dist <- claims %>%
  filter(Acc_Yr <= 2013) %>%           # mature AYs only
  distinct(claim_num, Acc_Yr, Rpt_Yr_Lag) %>%
  count(Rpt_Yr_Lag) %>%
  arrange(Rpt_Yr_Lag) %>%
  mutate(prop = n / sum(n),
         cumP = cumsum(prop))
cat("Rpt_Yr_Lag distribution (based on AY <= 2013 only):\n")
print(lag_dist)
n_obs_by_ay <- claims %>%
  distinct(claim_num, Acc_Yr) %>%
  count(Acc_Yr, name = "N_obs")
ibnr_counts <- n_obs_by_ay %>%
  mutate(
    thresh  = max_obs_cal - Acc_Yr + 1,
    P_obs   = sapply(thresh, function(t)
      sum(lag_dist$prop[lag_dist$Rpt_Yr_Lag <= t])),
    N_total = round(N_obs / P_obs),
    N_IBNR  = N_total - N_obs
  )
cat("\nExpected IBNR claims by AY:\n")
print(ibnr_counts %>% select(Acc_Yr, N_obs, thresh, P_obs, N_total, N_IBNR))
cat(sprintf("Total estimated pure IBNR claims: %d\n", sum(ibnr_counts$N_IBNR)))

# Generate synthetic IBNR claims
set.seed(42)
ibnr_synthetic <- ibnr_counts %>%
  filter(N_IBNR > 0) %>%
  rowwise() %>%
  reframe({
    ay  <- Acc_Yr
    n   <- N_IBNR
    thr <- thresh
    lag_pool <- lag_dist %>% filter(Rpt_Yr_Lag > thr)
    if (nrow(lag_pool) == 0 || n == 0) {
      tibble(claim_id = character(0), Acc_Yr = integer(0),
             Rpt_Yr_Lag = integer(0))
    } else {
      sampled_lags <- sample(lag_pool$Rpt_Yr_Lag,
                             size = n, replace = TRUE,
                             prob = lag_pool$prop / sum(lag_pool$prop))
      tibble(claim_id   = paste0("IBNR_", ay, "_", seq_len(n)),
             Acc_Yr     = as.integer(ay),
             Rpt_Yr_Lag = as.integer(sampled_lags))
    }
  })
cat(sprintf("Synthetic IBNR claims generated: %d\n", nrow(ibnr_synthetic)))

# Project synthetic IBNR claims with the same RF
# dev starts at Rpt_Yr_Lag - 1 so first d_next = Rpt_Yr_Lag
state_ibnr <- ibnr_synthetic %>%
  transmute(claim_num = claim_id, Acc_Yr, Rpt_Yr_Lag,
            dev       = as.integer(Rpt_Yr_Lag - 1L),
            cum       = 0,
            prob_open = 1)
reserves_ibnr <- tibble(Acc_Yr = ay_range, R_hat_IBNR = 0)
for (step in 1:max_dev) {
  state_ibnr <- state_ibnr %>% filter(dev < max_dev)
  if (nrow(state_ibnr) == 0) break
  d_next  <- state_ibnr$dev + 1
  cy_next <- state_ibnr$Acc_Yr + d_next - 1
  gen_idx <- infl_full$Accum_Infl_Index_Mid[match(cy_next, infl_full$Cal_Yr)]
  gen_idx[is.na(gen_idx)] <- tail(infl_full$Accum_Infl_Index_Mid, 1)
  X_ibnr <- data.frame(
    Acc_Yr          = state_ibnr$Acc_Yr,
    Dev_Yr          = d_next,
    Rpt_Yr_Lag      = state_ibnr$Rpt_Yr_Lag,
    has_paid_before = as.integer(state_ibnr$cum > 0),
    log_lag_cum     = log1p(state_ibnr$cum),
    dev_ratio       = d_next / max_dev
  )[, feats]
  p_pay_i   <- predict(rf_freq,  X_ibnr, type = "prob")[, "1"]
  log_sev_i <- predict(rf_sev,   X_ibnr)
  p_close_i <- predict(rf_close, X_ibnr, type = "prob")[, "1"]
  sev_nom_i <- exp(log_sev_i + sigma2 / 2) * gen_idx
  exp_pay_i <- state_ibnr$prob_open * p_pay_i * sev_nom_i
  tmp_i <- tibble(Acc_Yr = state_ibnr$Acc_Yr, pay = exp_pay_i) %>%
    group_by(Acc_Yr) %>% summarise(pay = sum(pay), .groups = "drop")
  reserves_ibnr <- reserves_ibnr %>%
    left_join(tmp_i, by = "Acc_Yr") %>%
    mutate(R_hat_IBNR = R_hat_IBNR + replace_na(pay, 0)) %>%
    select(-pay)
  
  # keep cum in nominal units (consistent with training)
  state_ibnr$cum       <- state_ibnr$cum + exp_pay_i
  state_ibnr$prob_open <- state_ibnr$prob_open * (1 - p_close_i)
  state_ibnr$dev       <- d_next
}
cat("\n=== Pure IBNR reserve ===\n")
print(reserves_ibnr %>% filter(R_hat_IBNR > 0))
cat(sprintf("Total pure IBNR: %s USD\n",
            format(round(sum(reserves_ibnr$R_hat_IBNR)), big.mark = ",")))

# Combined RBNS + pure IBNR
reserves_rf_total <- reserves_ml %>%
  rename(R_RBNS = R_hat_RF) %>%
  left_join(reserves_ibnr, by = "Acc_Yr") %>%
  mutate(R_hat_IBNR = replace_na(R_hat_IBNR, 0),
         R_RF_total = R_RBNS + R_hat_IBNR)
cat(sprintf("\nRF RBNS:       %s USD\n",
            format(round(sum(reserves_rf_total$R_RBNS)),      big.mark = ",")))
cat(sprintf("RF pure IBNR:  %s USD\n",
            format(round(sum(reserves_rf_total$R_hat_IBNR)),  big.mark = ",")))
cat(sprintf("RF TOTAL:      %s USD\n",
            format(round(sum(reserves_rf_total$R_RF_total)),  big.mark = ",")))
p_rbns_ibnr <- reserves_rf_total %>%
  select(Acc_Yr, `RBNS (already reported)` = R_RBNS,
         `Pure IBNR (not yet reported)` = R_hat_IBNR) %>%
  pivot_longer(-Acc_Yr, names_to = "Component", values_to = "Reserve") %>%
  ggplot(aes(factor(Acc_Yr), Reserve / 1e6, fill = Component)) +
  geom_col() +
  scale_fill_manual(values = c("RBNS (already reported)"   = "#e67e22",
                               "Pure IBNR (not yet reported)" = "#c0392b")) +
  scale_y_continuous(labels = comma) +
  labs(title = "RF reserve: RBNS + pure IBNR by accident year",
       x = "Accident Year", y = "Reserve (USD millions)") +
  theme(legend.position = "bottom")
ggsave("figs/rbns_ibnr.pdf", p_rbns_ibnr, width = 7, height = 4)
print(p_rbns_ibnr)
```

### Backtest Validation Framework
We use the same cut-off years for both models. For each CUT, only information 
available up to that calendar year is used.
Chain-Ladder:
* fitted on the truncated inflation-adjusted triangle
* projects future payments beyond CUT

Random Forest:
* trained on claim-level data observed up to CUT
* projects future payments for open claims

In both cases, predictions are compared with actual payments observed after CUT. 
We report Total_err_pct, i.e. the portfolio-level error: 
(total predicted - total actual) / total actual. This is the metric used in the graph.

```{r unified-backtest-validation, echo=TRUE}
# We report Total_err_pct:
# 100 * (total predicted - total actual) / total actual
# this is the portfolio-level error used in the graph.
cut_years <- 2016:2020

# Inflation-adjusted Chain-Ladder backtest
run_backtest_infl_cl <- function(CUT) {
  ay_train <- ay_range[ay_range <= CUT]
  triangle_cut <- expand_grid(Acc_Yr = ay_train, Dev_Yr = 1:max_dev) %>%
    mutate(
      Cal_Yr = Acc_Yr + Dev_Yr - 1,
      obs    = Cal_Yr <= CUT
    ) %>%
    left_join(triangle_incr, by = c("Acc_Yr", "Dev_Yr")) %>%
    mutate(S = ifelse(obs, replace_na(S, 0), NA_real_))
  triangle_defl_cut <- triangle_cut %>%
    left_join(infl_tbl, by = "Cal_Yr") %>%
    mutate(S_def = ifelse(obs, S / Accum_Infl_Index_Mid, NA_real_)) %>%
    group_by(Acc_Yr) %>%
    arrange(Dev_Yr, .by_group = TRUE) %>%
    mutate(C_def = ifelse(obs, cumsum(replace_na(S_def, 0)), NA_real_)) %>%
    ungroup()
  dev_factors_def_cut <- map_dfr(1:(max_dev - 1), function(k) {
    tmp <- triangle_defl_cut %>%
      filter(Dev_Yr %in% c(k, k + 1), obs) %>%
      select(Acc_Yr, Dev_Yr, C_def) %>%
      pivot_wider(
        names_from = Dev_Yr,
        values_from = C_def,
        names_prefix = "d"
      )
    col_k  <- paste0("d", k)
    col_k1 <- paste0("d", k + 1)
    if (!col_k %in% names(tmp) || !col_k1 %in% names(tmp)) {
      return(tibble(k = k, f_k = 1))
    }
    tmp <- tmp %>%
      filter(
        !is.na(.data[[col_k]]),
        !is.na(.data[[col_k1]]),
        .data[[col_k]] > 0
      )
    if (nrow(tmp) == 0) return(tibble(k = k, f_k = 1))
    tibble(
      k = k,
      f_k = sum(tmp[[col_k1]]) / sum(tmp[[col_k]])
    )
  })
  dev_factors_def_cut$f_k[dev_factors_def_cut$k == max_dev - 1] <- 1
  rect_def_cut <- triangle_defl_cut %>%
    select(Acc_Yr, Dev_Yr, Cal_Yr, obs, C_def) %>%
    arrange(Acc_Yr, Dev_Yr)
  for (ay in ay_train) {
    for (k in 1:(max_dev - 1)) {
      ic  <- which(rect_def_cut$Acc_Yr == ay & rect_def_cut$Dev_Yr == k)
      in_ <- which(rect_def_cut$Acc_Yr == ay & rect_def_cut$Dev_Yr == (k + 1))
      if (
        length(ic) > 0 && length(in_) > 0 &&
        !is.na(rect_def_cut$C_def[ic]) &&
        is.na(rect_def_cut$C_def[in_])
      ) {
        rect_def_cut$C_def[in_] <- rect_def_cut$C_def[ic] *
          dev_factors_def_cut$f_k[dev_factors_def_cut$k == k]
      }
    }
  }
  pred_window <- rect_def_cut %>%
    group_by(Acc_Yr) %>%
    arrange(Dev_Yr, .by_group = TRUE) %>%
    mutate(S_def_proj = C_def - lag(C_def, default = 0)) %>%
    ungroup() %>%
    filter(!obs, Cal_Yr > CUT, Cal_Yr <= max_obs_cal, S_def_proj > 0) %>%
    left_join(infl_tbl, by = "Cal_Yr") %>%
    mutate(S_proj = S_def_proj * Accum_Infl_Index_Mid) %>%
    group_by(Acc_Yr) %>%
    summarise(R_pred = sum(S_proj, na.rm = TRUE), .groups = "drop")
  actual_window <- claims %>%
    filter(Acc_Yr %in% ay_train, Cal_Yr > CUT, Cal_Yr <= max_obs_cal) %>%
    group_by(Acc_Yr) %>%
    summarise(R_actual = sum(Incr_Payment, na.rm = TRUE), .groups = "drop")
  pred_window %>%
    full_join(actual_window, by = "Acc_Yr") %>%
    mutate(
      R_pred    = replace_na(R_pred, 0),
      R_actual  = replace_na(R_actual, 0),
      Error     = R_pred - R_actual,
      Rel_Error = ifelse(R_actual > 0, Error / R_actual, NA_real_),
      CUT       = CUT
    )
}
bt_infl_all <- map_dfr(cut_years, run_backtest_infl_cl)
bt_infl_summary <- bt_infl_all %>%
  group_by(CUT) %>%
  summarise(
    n_AY           = n(),
    Total_pred_M   = sum(R_pred) / 1e6,
    Total_actual_M = sum(R_actual) / 1e6,
    Total_err_pct  = 100 * (sum(R_pred) - sum(R_actual)) /
      max(sum(R_actual), 1),
    .groups = "drop"
  )

# Random Forest backtest
run_rf_backtest <- function(CUT) {
  set.seed(1000 + CUT)
  claims_rf_tv <- claims_rf %>%
    mutate(in_train = Cal_Yr <= CUT)
  train_cv <- claims_rf_tv %>%
    filter(in_train, lag_open == 1)
  freq_tab <- table(as.factor(train_cv$has_payment))
  ss_freq  <- min(10000, floor(0.9 * min(freq_tab)))
  rf_freq_cv <- randomForest(
    x = train_cv[, feats],
    y = as.factor(train_cv$has_payment),
    ntree = 200,
    mtry = 3,
    maxnodes = 512,
    sampsize = c("0" = ss_freq, "1" = ss_freq)
  )
  sev_tr_cv <- train_cv %>%
    filter(has_payment == 1, !is.na(log_payment_defl))
  rf_sev_cv <- randomForest(
    x = sev_tr_cv[, feats],
    y = sev_tr_cv$log_payment_defl,
    ntree = 200,
    mtry = 3,
    maxnodes = 512
  )
  sigma2_cv <- mean((rf_sev_cv$y - rf_sev_cv$predicted)^2, na.rm = TRUE)
  close_tab <- table(as.factor(train_cv$closed_this_yr))
  ss_close  <- min(10000, floor(0.9 * min(close_tab)))
  rf_close_cv <- randomForest(
    x = train_cv[, feats],
    y = as.factor(train_cv$closed_this_yr),
    ntree = 200,
    mtry = 3,
    maxnodes = 512,
    sampsize = c("0" = ss_close, "1" = ss_close)
  )
  state_cv <- claims_rf_tv %>%
    filter(in_train) %>%
    group_by(claim_num) %>%
    filter(Dev_Yr == max(Dev_Yr)) %>%
    ungroup() %>%
    filter(open_ind == 1, Dev_Yr < max_dev) %>%
    transmute(
      claim_num,
      Acc_Yr,
      Rpt_Yr_Lag,
      dev = Dev_Yr,
      cum = Acc_Incr_Payment,
      prob_open = 1
    )
  proj_cv <- tibble(
    claim_num = character(),
    Acc_Yr = integer(),
    Cal_Yr_proj = integer(),
    pay_pred = double()
  )
  state_loop <- state_cv
  for (step in 1:max_dev) {
    state_loop <- state_loop %>%
      filter(dev < max_dev)
    if (nrow(state_loop) == 0) break
    d_next  <- state_loop$dev + 1
    cy_next <- state_loop$Acc_Yr + d_next - 1
    gen_idx <- infl_full$Accum_Infl_Index_Mid[
      match(cy_next, infl_full$Cal_Yr)
    ]
    gen_idx[is.na(gen_idx)] <- tail(infl_full$Accum_Infl_Index_Mid, 1)
    X_cv <- data.frame(
      Acc_Yr          = state_loop$Acc_Yr,
      Dev_Yr          = d_next,
      Rpt_Yr_Lag      = state_loop$Rpt_Yr_Lag,
      has_paid_before = as.integer(state_loop$cum > 0),
      log_lag_cum     = log1p(state_loop$cum),
      dev_ratio       = d_next / max_dev
    )[, feats]
    p_pay_cv   <- predict(rf_freq_cv, X_cv, type = "prob")[, "1"]
    log_sev_cv <- predict(rf_sev_cv, X_cv)
    p_close_cv <- predict(rf_close_cv, X_cv, type = "prob")[, "1"]
    sev_nom_cv <- exp(log_sev_cv + sigma2_cv / 2) * gen_idx
    exp_pay_cv <- state_loop$prob_open * p_pay_cv * sev_nom_cv
    proj_cv <- bind_rows(
      proj_cv,
      tibble(
        claim_num   = state_loop$claim_num,
        Acc_Yr      = state_loop$Acc_Yr,
        Cal_Yr_proj = cy_next,
        pay_pred    = exp_pay_cv
      )
    )
    state_loop$cum       <- state_loop$cum + exp_pay_cv
    state_loop$prob_open <- state_loop$prob_open * (1 - p_close_cv)
    state_loop$dev       <- d_next
  }
  real_pay <- claims %>%
    filter(
      claim_num %in% state_cv$claim_num,
      Cal_Yr > CUT,
      Cal_Yr <= max_obs_cal
    ) %>%
    group_by(Acc_Yr) %>%
    summarise(R_actual = sum(Incr_Payment, na.rm = TRUE), .groups = "drop")
  pred_pay <- proj_cv %>%
    filter(Cal_Yr_proj > CUT, Cal_Yr_proj <= max_obs_cal) %>%
    group_by(Acc_Yr) %>%
    summarise(R_pred = sum(pay_pred), .groups = "drop")
  pred_pay %>%
    full_join(real_pay, by = "Acc_Yr") %>%
    mutate(
      R_pred    = replace_na(R_pred, 0),
      R_actual  = replace_na(R_actual, 0),
      Error     = R_pred - R_actual,
      Rel_Error = ifelse(R_actual > 0, Error / R_actual, NA_real_),
      CUT       = CUT
    )
}
bt_rf_all <- map_dfr(cut_years, run_rf_backtest)
bt_rf_summary <- bt_rf_all %>%
  group_by(CUT) %>%
  summarise(
    n_AY           = n(),
    Total_pred_M   = sum(R_pred) / 1e6,
    Total_actual_M = sum(R_actual) / 1e6,
    Total_err_pct  = 100 * (sum(R_pred) - sum(R_actual)) /
      max(sum(R_actual), 1),
    .groups = "drop"
  )

# comparison
bt_compare <- bind_rows(
  bt_infl_summary %>% mutate(Method = "CL Inflation-Adjusted"),
  bt_rf_summary   %>% mutate(Method = "Random Forest")
)
bt_compare_table <- bt_compare %>%
  select(CUT, Method, Total_pred_M, Total_actual_M, Total_err_pct) %>%
  mutate(
    Total_pred_M   = round(Total_pred_M, 1),
    Total_actual_M = round(Total_actual_M, 1),
    Total_err_pct  = round(Total_err_pct, 1)
  ) %>%
  arrange(CUT, Method)
cat("\n=== Portfolio-level total error by cut-off year ===\n")
print(bt_compare_table)
bt_compare_wide <- bt_compare %>%
  select(CUT, Method, Total_err_pct) %>%
  pivot_wider(
    names_from = Method,
    values_from = Total_err_pct
  ) %>%
  mutate(across(where(is.numeric), ~ round(., 1))) %>%
  arrange(CUT)
cat("\n=== Total error (%) used in the graph ===\n")
print(bt_compare_wide)
p_bt_compare <- bt_compare %>%
  ggplot(aes(x = factor(CUT), y = Total_err_pct, fill = Method)) +
  geom_col(position = "dodge") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = c(
    "CL Inflation-Adjusted" = "#2ecc71",
    "Random Forest"        = "#e74c3c"
  )) +
  labs(
    title = "Backtest: inflation-adjusted CL vs RF — total error by cut-off year",
    subtitle = "Negative = underestimation, positive = overestimation",
    x = "Cut-off year",
    y = "Total error (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
ggsave(filename = "figs/backtest_infl_cl_vs_rf.png", plot = p_bt_compare,
       width = 8, height = 4, dpi = 300)
print(p_bt_compare)
```

### Cox PH Correction for Right-Censoring

```{r cox right-censoring, echo=TRUE}
# Survival dataset (one row per claim)
# event = 1 if the claim was observed to close, 0 if still open
# at its last observed Dev_Yr (right-censored)
surv_data <- claims_rf %>%
  group_by(claim_num) %>%
  summarise(
    Acc_Yr         = first(Acc_Yr),
    Rpt_Yr_Lag     = first(Rpt_Yr_Lag),
    time_obs       = as.numeric(max(Dev_Yr) - min(Dev_Yr) + 1),
    event          = as.integer(any(open_ind == 0)),
    log_total_paid = log1p(max(Acc_Incr_Payment)),
    has_paid       = as.integer(max(Acc_Incr_Payment) > 0),
    .groups        = "drop"
  ) %>%
  filter(time_obs >= 1)   # drop degenerate cases 
cat(sprintf("Survival dataset: %d claims\n", nrow(surv_data)))
cat(sprintf("  Closed   (event=1): %d  (%.1f%%)\n",
            sum(surv_data$event == 1), 100 * mean(surv_data$event == 1)))
cat(sprintf("  Censored (event=0): %d  (%.1f%%)\n",
            sum(surv_data$event == 0), 100 * mean(surv_data$event == 0)))
cat("\nCensoring rate by Acc_Yr:\n")
print(surv_data %>%
        group_by(Acc_Yr) %>%
        summarise(n_claims   = n(),
                  n_censored = sum(event == 0),
                  pct_cens   = round(100 * mean(event == 0), 1),
                  .groups    = "drop"))

# Kaplan-Meier (unconditional baseline)
km_fit  <- survfit(Surv(time_obs, event) ~ 1, data = surv_data)
km_summ <- summary(km_fit, times = 1:max_dev)

cat("\nKaplan-Meier S(t) = P(claim still open at Dev_Yr t):\n")
print(data.frame(Dev_Yr  = km_summ$time,
                 S_t     = round(km_summ$surv, 3),
                 n_risk  = km_summ$n.risk,
                 n_event = km_summ$n.event))
p_km <- data.frame(t  = km_summ$time, S  = km_summ$surv,
                   lo = km_summ$lower, hi = km_summ$upper) %>%
  ggplot(aes(t, S)) +
  geom_step(color = "#1b4f72", linewidth = 1) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.2, fill = "#1b4f72") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey50") +
  scale_x_continuous(breaks = 1:max_dev) +
  labs(title = "Kaplan-Meier survival curve",
       subtitle = "S(t) = P(claim still open at Dev_Yr t) — band = 95% CI",
       x = "Development Year", y = "P(claim open)")
ggsave("figs/kaplan_meier.pdf", p_km, width = 6, height = 4)
print(p_km)

# Cox PH model
# Acc_Yr as continuous covariate to share one baseline
# hazard across all AYs — avoids flat curves in recent AYs with few events.
cox_fit <- coxph(
  Surv(time_obs, event) ~
    Acc_Yr + Rpt_Yr_Lag + has_paid + log_total_paid,
  data = surv_data,
  ties = "efron"
)
cat("\n=== Cox PH summary ===\n")
print(summary(cox_fit))
cat(sprintf("C-index: %.4f\n\n", summary(cox_fit)$concordance["C"]))

# get_survival_matrix(): for each row of newdata, compute S(t) at
# each integer t = 1, ..., max_dev. Returns matrix [n_claims x (max_dev+1)]
# where column 1 = S(0) = 1 and column k+1 = S(k)
get_survival_matrix <- function(newdata, times = 1:max_dev) {
  n  <- nrow(newdata)
  sf <- survfit(cox_fit, newdata = newdata, se.fit = FALSE)
  t_raw <- sf$time
  s_raw <- sf$surv
  if (!is.matrix(s_raw)) s_raw <- matrix(s_raw, ncol = 1)
  S_mat <- matrix(NA_real_, nrow = n, ncol = length(times))
  for (j in seq_len(n)) {
    S_mat[j, ] <- approx(t_raw, s_raw[, j], xout = times,
                         method = "constant", rule = 2, f = 0)$y
  }
  cbind(1, S_mat)  
}

# conditional_p_close(): 1 - S(d)/S(d-1), clamped to [0, 1]
conditional_p_close <- function(S_with_t0, d_vec) {
  n      <- nrow(S_with_t0)
  d_safe <- pmin(as.integer(d_vec), max_dev)
  S_prev <- S_with_t0[cbind(seq_len(n), d_safe)]
  S_curr <- S_with_t0[cbind(seq_len(n), d_safe + 1L)]
  pmax(0, pmin(1, 1 - S_curr / pmax(S_prev, 1e-10)))
}

# RBNS projection with Cox closure
open_rbns <- claims_rf %>%
  group_by(claim_num) %>%
  filter(Dev_Yr == max(Dev_Yr)) %>% ungroup() %>%
  filter(open_ind == 1, Dev_Yr < max_dev)
cat(sprintf("Open RBNS claims to project: %d\n", nrow(open_rbns)))
cox_nd_rbns <- open_rbns %>%
  transmute(Acc_Yr, Rpt_Yr_Lag,
            has_paid       = as.integer(Acc_Incr_Payment > 0),
            log_total_paid = log1p(Acc_Incr_Payment))
S_rbns <- get_survival_matrix(cox_nd_rbns)
state_cox <- open_rbns %>%
  transmute(claim_num, Acc_Yr, Rpt_Yr_Lag,
            dev = Dev_Yr, cum = Acc_Incr_Payment,
            prob_open = 1, row_idx = row_number())
reserves_cox_rbns <- tibble(Acc_Yr = ay_range, R_Cox_RBNS = 0)
for (step in seq_len(max_dev)) {
  state_cox <- state_cox %>% filter(dev < max_dev)
  if (nrow(state_cox) == 0) break
  d_next  <- state_cox$dev + 1
  cy_next <- state_cox$Acc_Yr + d_next - 1
  gen_idx <- infl_full$Accum_Infl_Index_Mid[match(cy_next, infl_full$Cal_Yr)]
  gen_idx[is.na(gen_idx)] <- tail(infl_full$Accum_Infl_Index_Mid, 1)
  X <- data.frame(
    Acc_Yr          = state_cox$Acc_Yr,
    Dev_Yr          = d_next,
    Rpt_Yr_Lag      = state_cox$Rpt_Yr_Lag,
    has_paid_before = as.integer(state_cox$cum > 0),
    log_lag_cum     = log1p(state_cox$cum),
    dev_ratio       = d_next / max_dev
  )[, feats]
  p_pay   <- predict(rf_freq, X, type = "prob")[, "1"]
  log_sev <- predict(rf_sev,  X)
  sev_nom <- exp(log_sev + sigma2 / 2) * gen_idx
  
  # Cox is built on elapsed time (max(Dev_Yr) - min(Dev_Yr) + 1)
  # so the survival lookup must also use elapsed time, not absolute Dev_Yr
  # Elapsed at end of step k is: d_next - Rpt_Yr_Lag + 1
  elapsed_next <- d_next - state_cox$Rpt_Yr_Lag + 1
  elapsed_next <- pmax(elapsed_next, 1L)
  p_close <- conditional_p_close(
    S_rbns[state_cox$row_idx, , drop = FALSE],
    d_vec = elapsed_next
  )
  exp_pay <- state_cox$prob_open * p_pay * sev_nom
  tmp <- tibble(Acc_Yr = state_cox$Acc_Yr, pay = exp_pay) %>%
    group_by(Acc_Yr) %>% summarise(pay = sum(pay), .groups = "drop")
  reserves_cox_rbns <- reserves_cox_rbns %>%
    left_join(tmp, by = "Acc_Yr") %>%
    mutate(R_Cox_RBNS = R_Cox_RBNS + replace_na(pay, 0)) %>%
    select(-pay)
  state_cox$cum       <- state_cox$cum + exp_pay
  state_cox$prob_open <- state_cox$prob_open * (1 - p_close)
  state_cox$dev       <- d_next
}
cat("=== RBNS reserve (Cox closure) ===\n")
print(reserves_cox_rbns)
cat(sprintf("Total RBNS (Cox): %s USD\n\n",
            format(round(sum(reserves_cox_rbns$R_Cox_RBNS)), big.mark = ",")))

# Final RF total: RBNS(Cox) + IBNR(rf_close)
# IBNR uses rf_close (not Cox): synthetic claims have no observed history
# so right-censoring bias does not apply
# rf_close is unbiased at low Dev_Yr where IBNR claims start
reserves_cox_total <- reserves_cox_rbns %>%
  left_join(reserves_ibnr, by = "Acc_Yr") %>% 
  mutate(
    R_hat_IBNR  = replace_na(R_hat_IBNR, 0),
    R_Cox_total = R_Cox_RBNS + R_hat_IBNR
  )
cat("=== RF FINAL: RBNS(Cox) + IBNR(rf_close) ===\n")
print(reserves_cox_total)
cat(sprintf("RF RBNS  (Cox closure):      %s USD\n",
            format(round(sum(reserves_cox_total$R_Cox_RBNS)),  big.mark = ",")))
cat(sprintf("RF IBNR  (rf_close, no Cox): %s USD\n",
            format(round(sum(reserves_cox_total$R_hat_IBNR)),  big.mark = ",")))
cat(sprintf("RF TOTAL:                    %s USD\n",
            format(round(sum(reserves_cox_total$R_Cox_total)), big.mark = ",")))
```

### Full Comparison Plot

We compare the different methods used to caculate the reserve.

```{r methods, echo=TRUE}
compare_all <- reserve_infl %>%
  select(Acc_Yr, R_CL = R_hat_infl) %>%
  left_join(reserves_scenarios %>% select(Acc_Yr, R_CL_mod = R_moderate), by = "Acc_Yr") %>%
  left_join(reserves_ml        %>% rename(R_RF_orig = R_hat_RF),          by = "Acc_Yr") %>%
  left_join(reserves_rf_total  %>% select(Acc_Yr, R_RF_IBNR = R_RF_total), by = "Acc_Yr") %>%
  left_join(reserves_cox_total %>% select(Acc_Yr, R_Cox_total),           by = "Acc_Yr") %>%
  mutate(across(where(is.numeric), ~ replace_na(., 0)))

p_all <- compare_all %>%
  select(Acc_Yr,
         "1. CL infl-adjusted"          = R_CL,
         "2. CL moderate super-infl"    = R_CL_mod,
         "3. RF original"               = R_RF_orig,
         "4. RF + IBNR"                 = R_RF_IBNR,
         "5. RF Cox + IBNR"             = R_Cox_total) %>%
  pivot_longer(-Acc_Yr, names_to = "Method", values_to = "Reserve") %>%
  ggplot(aes(factor(Acc_Yr), Reserve / 1e6, fill = Method)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("1. CL infl-adjusted"       = "#1b4f72",
                               "2. CL moderate super-infl" = "#8e44ad",
                               "3. RF original"            = "#e67e22",
                               "4. RF + IBNR"              = "#c0392b",
                               "5. RF Cox + IBNR"          = "#27ae60")) +
  scale_y_continuous(labels = comma) +
  labs(title = "Reserve estimates — all methods by Accident Year",
       x = "Accident Year", y = "Reserve (USD millions)") +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 8))
ggsave("figs/all_methods.pdf", p_all, width = 8, height = 4)
print(p_all)
```

## Bootstrap for Chain Ladder

```{r bootstrap CL inflation-adjusted, echo=TRUE}
B <- 100   # bootstrap replicates

# (1) Use deflated development factors and deflated triangle
f_vec        <- f_def$f_def
names(f_vec) <- as.character(f_def$k)

# Back-fit the fitted DEFLATED cumulative & incremental triangle
# Convention: on the latest observed diagonal, fitted = observed
# Earlier cumulatives are recovered via hat_C_{i,j-1} = hat_C_{i,j} / f_{j-1}
fitted_tri <- tri_defl %>%
  filter(obs) %>%
  select(Acc_Yr, Dev_Yr, Cal_Yr, C_def) %>%
  rename(C = C_def) %>%               # uniformiser le nom pour le reste du code
  arrange(Acc_Yr, Dev_Yr)
fitted_tri$hat_C <- NA_real_
for (ay in unique(fitted_tri$Acc_Yr)) {
  rows  <- which(fitted_tri$Acc_Yr == ay)
  ks    <- fitted_tri$Dev_Yr[rows]
  k_max <- max(ks)
  fitted_tri$hat_C[rows[ks == k_max]] <- fitted_tri$C[rows[ks == k_max]]
  if (k_max > 1) for (k in seq(k_max - 1, 1)) {
    fitted_tri$hat_C[rows[ks == k]] <-
      fitted_tri$hat_C[rows[ks == k + 1]] / f_vec[as.character(k)]
  }
}
fitted_tri <- fitted_tri %>%
  group_by(Acc_Yr) %>% arrange(Dev_Yr, .by_group = TRUE) %>%
  mutate(hat_S = hat_C - lag(hat_C, default = 0),
         S     = C     - lag(C,     default = 0)) %>%
  ungroup()

# Pearson residuals on DEFLATED scale + degrees-of-freedom adjustment
res_pool_df <- fitted_tri %>%
  filter(hat_S > 0) %>%
  mutate(pearson = (S - hat_S) / sqrt(hat_S))
N <- nrow(res_pool_df)
p <- 2 * max_dev - 1
res_pool <- res_pool_df$pearson * sqrt(N / (N - p))
cat(sprintf("Residual pool (deflated): N=%d cells, p=%d params, scale=%.4f\n",
            N, p, sqrt(N / (N - p))))

# Bootstrap loop
boot_total <- numeric(B)
boot_by_ay <- matrix(NA_real_, nrow = B, ncol = length(ay_range),
                     dimnames = list(NULL, as.character(ay_range)))
for (b in seq_len(B)) {

  # (a) Resample residuals -> pseudo-S (deflated)
  pseudo <- fitted_tri %>%
    mutate(r_star = sample(res_pool, n(), replace = TRUE),
           S_star = ifelse(hat_S > 0,
                           hat_S + r_star * sqrt(pmax(hat_S, 0)),
                           hat_S),
           S_star = pmax(S_star, 0))

  # (b) Re-cumulate the pseudo upper triangle (deflated)
  cum_star <- pseudo %>%
    group_by(Acc_Yr) %>% arrange(Dev_Yr, .by_group = TRUE) %>%
    mutate(C_star = cumsum(S_star)) %>% ungroup()

  # (c) Re-estimate volume-weighted f* on the pseudo triangle
  f_star <- map_dfr(1:(max_dev - 1), function(k) {
    tmp <- cum_star %>%
      filter(Dev_Yr %in% c(k, k + 1)) %>%
      select(Acc_Yr, Dev_Yr, C_star) %>%
      pivot_wider(names_from = Dev_Yr, values_from = C_star,
                  names_prefix = "d") %>%
      filter(!is.na(.data[[paste0("d", k)]]),
             !is.na(.data[[paste0("d", k + 1)]]),
             .data[[paste0("d", k)]] > 0)
    if (nrow(tmp) == 0 || sum(tmp[[paste0("d", k)]]) == 0)
      return(tibble(k = k, f = 1))
    tibble(k = k,
           f = sum(tmp[[paste0("d", k + 1)]]) /
               sum(tmp[[paste0("d", k)]]))
  })
  f_star$f[f_star$k == max_dev - 1] <- 1

  # (d) Project the full DEFLATED rectangle one step at a time
  rect_b <- expand_grid(Acc_Yr = ay_range, Dev_Yr = 1:max_dev) %>%
    mutate(Cal_Yr = Acc_Yr + Dev_Yr - 1,
           obs    = Cal_Yr <= max_cal) %>%
    left_join(cum_star %>% select(Acc_Yr, Dev_Yr, C_star),
              by = c("Acc_Yr", "Dev_Yr")) %>%
    arrange(Acc_Yr, Dev_Yr)

  for (ay in ay_range) for (k in 1:(max_dev - 1)) {
    ic  <- which(rect_b$Acc_Yr == ay & rect_b$Dev_Yr == k)
    in_ <- which(rect_b$Acc_Yr == ay & rect_b$Dev_Yr == (k + 1))
    if (length(ic) > 0 && length(in_) > 0 &&
        !is.na(rect_b$C_star[ic]) && is.na(rect_b$C_star[in_]))
      rect_b$C_star[in_] <- rect_b$C_star[ic] *
        f_star$f[f_star$k == k]
  }

  # (e) Deflated future incrementals -> RE-INFLATE -> nominal reserve per AY
  res_b <- rect_b %>%
    group_by(Acc_Yr) %>% arrange(Dev_Yr, .by_group = TRUE) %>%
    mutate(S_def_proj = C_star - lag(C_star, default = 0)) %>%
    ungroup() %>%
    filter(!obs) %>%
    left_join(infl_full, by = "Cal_Yr") %>%
    mutate(S_nom_proj = S_def_proj * Accum_Infl_Index_Mid) %>%
    group_by(Acc_Yr) %>%
    summarise(R = sum(S_nom_proj, na.rm = TRUE), .groups = "drop")

  boot_total[b] <- sum(res_b$R)
  boot_by_ay[b, as.character(res_b$Acc_Yr)] <- res_b$R
}

# Percentile CIs — compare against the inflation-adjusted point estimate
ci_total <- quantile(boot_total, c(0.025, 0.5, 0.975))
R_point  <- sum(reserve_infl$R_hat_infl)
cat("\n=== Bootstrap CI for total INFLATION-ADJUSTED chain-ladder reserve ===\n")
cat(sprintf("  B replicates:        %d\n", B))
cat(sprintf("  CL point estimate:   %s USD\n",
            format(round(R_point),     big.mark = ",")))
cat(sprintf("  Bootstrap median:    %s USD\n",
            format(round(ci_total[2]), big.mark = ",")))
cat(sprintf("  Bootstrap mean:      %s USD\n",
            format(round(mean(boot_total)), big.mark = ",")))
cat(sprintf("  Bootstrap SE:        %s USD  (CV = %.1f%%)\n",
            format(round(sd(boot_total)),   big.mark = ","),
            100 * sd(boot_total) / mean(boot_total)))
cat(sprintf("  95%% percentile CI:  [%s ; %s] USD\n",
            format(round(ci_total[1]), big.mark = ","),
            format(round(ci_total[3]), big.mark = ",")))

# Per-AY CIs
ci_by_ay <- apply(boot_by_ay, 2, quantile,
                  probs = c(0.025, 0.5, 0.975), na.rm = TRUE) %>%
  t() %>% as.data.frame() %>%
  setNames(c("lo", "med", "hi")) %>%
  tibble::rownames_to_column("Acc_Yr") %>%
  mutate(Acc_Yr = as.integer(Acc_Yr)) %>%
  left_join(reserve_infl %>% rename(R_hat = R_hat_infl), by = "Acc_Yr") %>%
  select(Acc_Yr, R_hat, lo, med, hi)
cat("\nBootstrap CI by accident year (USD millions):\n")
print(ci_by_ay %>%
        mutate(across(c(R_hat, lo, med, hi), ~ round(. / 1e6, 2))))

# Plot the predictive distribution
p_boot <- tibble(R = boot_total) %>%
  ggplot(aes(R / 1e6)) +
  geom_histogram(bins = 15, fill = "steelblue", color = "white") +
  geom_vline(xintercept = R_point / 1e6, color = "red", linewidth = 1) +
  geom_vline(xintercept = ci_total[c(1, 3)] / 1e6,
             color = "darkgreen", linetype = 2, linewidth = 0.7) +
  scale_x_continuous(labels = comma) +
  labs(title    = sprintf("Bootstrap predictive distribution inflation-adjusted CL (B=%d)", B),
       subtitle = "red = inflation-adjusted CL point estimate; green dashed = 95% percentile CI",
       x = "Total reserve (USD millions)", y = "Count")
ggsave("figs/bootstrap_CL_inflation.pdf", p_boot, width = 7, height = 4)
print(p_boot)
```

## Bootstrap for Random Forest

```{r bootstrap RF, echo=TRUE}
# Two sources of uncertainty:
# 1. Sampling variability in historical claims 
# 2. Poisson uncertainty on N_IBNR            
FEATS    <- feats
MAX_DEV  <- max_dev
claims   <- claims_rf
result   <- reserves_rf_total %>% rename(R_TOTAL = R_RF_total)
get_infl <- function(cy) {
  i <- infl_full$Accum_Infl_Index_Mid[match(cy, infl_full$Cal_Yr)]
  i[is.na(i)] <- tail(infl_full$Accum_Infl_Index_Mid, 1)
  i
}
boot_one <- function(boot_claims, seed) {
  set.seed(seed)
  rf_freq_b <- randomForest(
    x = boot_claims[, FEATS], y = factor(boot_claims$has_payment),
    ntree = 500, mtry = 3, maxnodes = 512, nodesize = 20,
    classwt = c("0" = 1,
                "1" = sum(boot_claims$has_payment == 0) /
                  max(sum(boot_claims$has_payment == 1), 1))
  )
  sev_b <- boot_claims %>%
    filter(has_payment == 1, !is.na(log_payment_defl))
  rf_sev_b <- randomForest(
    x = sev_b[, FEATS], y = sev_b$log_payment_defl,
    ntree = 500, mtry = 2, maxnodes = 256, nodesize = 20
  )
  sigma2_b <- mean((sev_b$log_payment_defl -
                      predict(rf_sev_b, sev_b[, FEATS]))^2, na.rm = TRUE)
  rf_close_b <- randomForest(
    x = boot_claims[, FEATS], y = factor(boot_claims$closed_this_yr),
    ntree = 500, mtry = 3, maxnodes = 512, nodesize = 20,
    classwt = c("0" = 1,
                "1" = sum(boot_claims$closed_this_yr == 0) /
                  max(sum(boot_claims$closed_this_yr == 1), 1))
  )
  proj <- function(state) {
    total <- 0
    for (s in seq_len(MAX_DEV)) {
      state <- state %>% filter(dev < MAX_DEV); if (nrow(state) == 0) break
      d_next  <- state$dev + 1
      cy_next <- state$Acc_Yr + d_next - 1
      X <- data.frame(
        Acc_Yr          = state$Acc_Yr,
        Dev_Yr          = d_next,
        Rpt_Yr_Lag      = state$Rpt_Yr_Lag,
        has_paid_before = as.integer(state$cum > 0),
        log_lag_cum     = log1p(state$cum),
        dev_ratio       = d_next / MAX_DEV
      )[, FEATS]
      p_pay   <- predict(rf_freq_b,  X, type = "prob")[, "1"]
      log_sev <- predict(rf_sev_b,   X)
      p_close <- predict(rf_close_b, X, type = "prob")[, "1"]
      sev_nom <- exp(log_sev + sigma2_b / 2) * get_infl(cy_next)
      exp_pay <- state$prob_open * p_pay * sev_nom
      total   <- total + sum(exp_pay)
      state   <- state %>% mutate(
        cum       = cum + exp_pay,
        prob_open = prob_open * (1 - p_close),
        dev       = d_next
      )
    }
    total
  }
  state_rbns <- claims %>%
    group_by(claim_num) %>% slice_max(Dev_Yr) %>% ungroup() %>%
    filter(open_ind == 1, Dev_Yr < MAX_DEV) %>%
    transmute(claim_num, Acc_Yr, Rpt_Yr_Lag,
              dev = Dev_Yr, cum = Acc_Incr_Payment, prob_open = 1.0)
  R_rbns <- proj(state_rbns)
  ibnr_b <- ibnr_counts %>%
    rowwise() %>%
    mutate(N_IBNR_b = rpois(1, lambda = max(N_IBNR, 0))) %>%
    ungroup()
  ibnr_synth_b <- purrr::map_dfr(seq_len(nrow(ibnr_b)), function(i) {
    r <- ibnr_b[i, ]; if (r$N_IBNR_b == 0) return(tibble())
    pool <- lag_dist %>% filter(Rpt_Yr_Lag > r$thresh)
    if (nrow(pool) == 0) return(tibble())
    ll <- sample(pool$Rpt_Yr_Lag, size = r$N_IBNR_b, replace = TRUE,
                 prob = pool$prop / sum(pool$prop))
    tibble(claim_id   = paste0("IBNR_", r$Acc_Yr, "_", seq_len(r$N_IBNR_b)),
           Acc_Yr     = as.integer(r$Acc_Yr),
           Rpt_Yr_Lag = as.integer(ll))
  })
  state_ibnr <- ibnr_synth_b %>%
    transmute(claim_num = claim_id, Acc_Yr, Rpt_Yr_Lag,
              dev = as.integer(Rpt_Yr_Lag - 1L),
              cum = 0, prob_open = 1.0)
  R_ibnr <- if (nrow(state_ibnr) > 0) proj(state_ibnr) else 0
  c(RBNS = R_rbns, IBNR = R_ibnr, TOTAL = R_rbns + R_ibnr)
}

# Bootstrap loop
B <- 100
claim_ids <- unique(claims$claim_num)
cat(sprintf("\nRunning %d bootstrap replicates...\n", B))
boot_results <- matrix(NA, nrow = B, ncol = 3,
                       dimnames = list(NULL, c("RBNS", "IBNR", "TOTAL")))
t0 <- Sys.time()
for (b in seq_len(B)) {
  set.seed(1000 + b)
  sampled_ids <- sample(claim_ids, length(claim_ids), replace = TRUE)
  boot_claims <- tibble(claim_num = sampled_ids,
                        boot_id = seq_along(sampled_ids)) %>%
    inner_join(claims, by = "claim_num", relationship = "many-to-many") %>%
    mutate(claim_num = paste0(claim_num, "_b", boot_id)) %>%
    select(-boot_id)
  boot_results[b, ] <- tryCatch(
    boot_one(boot_claims, seed = 2000 + b),
    error = function(e) { cat("  iter", b, "failed:", e$message, "\n"); rep(NA, 3) }
  )
  cat(sprintf("  iter %2d/%d: TOTAL = $%s   elapsed: %.1f min\n",
              b, B,
              format(round(boot_results[b, "TOTAL"]), big.mark = ","),
              as.numeric(difftime(Sys.time(), t0, units = "mins"))))
}

# Summarize — 95% CI (2.5% / 97.5%)
boot_df <- as_tibble(boot_results) %>% filter(!is.na(TOTAL))

ci_fn <- function(x, p = c(0.025, 0.50, 0.975)) {
  setNames(quantile(x, p, na.rm = TRUE), c("p025", "p50", "p975"))
}
cat("\n══════════ BOOTSTRAP RESULTS ══════════\n")
cat(sprintf("Replicates that succeeded: %d / %d\n\n", nrow(boot_df), B))
for (col in c("RBNS", "IBNR", "TOTAL")) {
  q <- ci_fn(boot_df[[col]])
  cat(sprintf("%-5s : mean = $%s, 95%% CI = [$%s, $%s]\n",
              col,
              format(round(mean(boot_df[[col]])),  big.mark = ","),
              format(round(q["p025"]),              big.mark = ","),
              format(round(q["p975"]),              big.mark = ",")))
}
cat(sprintf("\nPoint estimate from main run: $%s\n",
            format(round(sum(result$R_TOTAL)), big.mark = ",")))

# Histogram (95% CI)
p_boot_rf <- ggplot(boot_df, aes(x = TOTAL / 1e6)) +
  geom_histogram(bins = 20, fill = "#534AB7", color = "white") +
  geom_vline(xintercept = sum(result$R_TOTAL) / 1e6,
             color = "#1D9E75", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = quantile(boot_df$TOTAL, c(0.025, 0.975)) / 1e6,
             color = "black", linetype = "dotted") +
  labs(title    = sprintf("Bootstrap distribution of total RF reserve (B = %d)", nrow(boot_df)),
       subtitle = sprintf("Point estimate $%.0fM, 95%% CI [$%.0fM, $%.0fM]",
                          sum(result$R_TOTAL)/1e6,
                          quantile(boot_df$TOTAL, 0.025)/1e6,
                          quantile(boot_df$TOTAL, 0.975)/1e6),
       x = "Reserve (USD millions)", y = "Frequency") +
  theme_minimal(base_size = 12)
ggsave("figs/bootstrap_RF.pdf", p_boot_rf, width = 7, height = 4)
print(p_boot_rf)
```

```{r bootstrap RF by AY, echo=TRUE}
FEATS    <- feats
MAX_DEV  <- max_dev
claims   <- claims_rf
result   <- reserves_rf_total %>% rename(R_TOTAL = R_RF_total)
get_infl <- function(cy) {
  i <- infl_full$Accum_Infl_Index_Mid[match(cy, infl_full$Cal_Yr)]
  i[is.na(i)] <- tail(infl_full$Accum_Infl_Index_Mid, 1)
  i
}
boot_one <- function(boot_claims, seed) {
  set.seed(seed)
  
  # training of the 3 RFs
  rf_freq_b <- randomForest(
    x = boot_claims[, FEATS], y = factor(boot_claims$has_payment),
    ntree = 500, mtry = 3, maxnodes = 512, nodesize = 20,
    classwt = c("0" = 1,
                "1" = sum(boot_claims$has_payment == 0) /
                  max(sum(boot_claims$has_payment == 1), 1))
  )
  sev_b <- boot_claims %>%
    filter(has_payment == 1, !is.na(log_payment_defl))
  rf_sev_b <- randomForest(
    x = sev_b[, FEATS], y = sev_b$log_payment_defl,
    ntree = 500, mtry = 2, maxnodes = 256, nodesize = 20
  )
  sigma2_b <- mean((sev_b$log_payment_defl -
                      predict(rf_sev_b, sev_b[, FEATS]))^2, na.rm = TRUE)
  rf_close_b <- randomForest(
    x = boot_claims[, FEATS], y = factor(boot_claims$closed_this_yr),
    ntree = 500, mtry = 3, maxnodes = 512, nodesize = 20,
    classwt = c("0" = 1,
                "1" = sum(boot_claims$closed_this_yr == 0) /
                  max(sum(boot_claims$closed_this_yr == 1), 1))
  )
  
  proj <- function(state) {
    res_by_ay <- tibble(Acc_Yr = ay_range, R = 0)
    for (s in seq_len(MAX_DEV)) {
      state <- state %>% filter(dev < MAX_DEV); if (nrow(state) == 0) break
      d_next  <- state$dev + 1
      cy_next <- state$Acc_Yr + d_next - 1
      X <- data.frame(
        Acc_Yr          = state$Acc_Yr,
        Dev_Yr          = d_next,
        Rpt_Yr_Lag      = state$Rpt_Yr_Lag,
        has_paid_before = as.integer(state$cum > 0),
        log_lag_cum     = log1p(state$cum),
        dev_ratio       = d_next / MAX_DEV
      )[, FEATS]
      p_pay   <- predict(rf_freq_b,  X, type = "prob")[, "1"]
      log_sev <- predict(rf_sev_b,   X)
      p_close <- predict(rf_close_b, X, type = "prob")[, "1"]
      sev_nom <- exp(log_sev + sigma2_b / 2) * get_infl(cy_next)
      exp_pay <- state$prob_open * p_pay * sev_nom
      
      # accumulate by Acc_Yr
      tmp <- tibble(Acc_Yr = state$Acc_Yr, pay = exp_pay) %>%
        group_by(Acc_Yr) %>% summarise(pay = sum(pay), .groups = "drop")
      res_by_ay <- res_by_ay %>%
        left_join(tmp, by = "Acc_Yr") %>%
        mutate(R = R + replace_na(pay, 0)) %>%
        select(-pay)
      state <- state %>% mutate(
        cum       = cum + exp_pay,
        prob_open = prob_open * (1 - p_close),
        dev       = d_next
      )
    }
    res_by_ay
  }
  
  # RBNS
  state_rbns <- claims %>%
    group_by(claim_num) %>% slice_max(Dev_Yr) %>% ungroup() %>%
    filter(open_ind == 1, Dev_Yr < MAX_DEV) %>%
    transmute(claim_num, Acc_Yr, Rpt_Yr_Lag,
              dev = Dev_Yr, cum = Acc_Incr_Payment, prob_open = 1.0)
  R_rbns_ay <- proj(state_rbns)
  
  # IBNR (identical)
  ibnr_b <- ibnr_counts %>%
    rowwise() %>%
    mutate(N_IBNR_b = rpois(1, lambda = max(N_IBNR, 0))) %>%
    ungroup()
  ibnr_synth_b <- purrr::map_dfr(seq_len(nrow(ibnr_b)), function(i) {
    r <- ibnr_b[i, ]; if (r$N_IBNR_b == 0) return(tibble())
    pool <- lag_dist %>% filter(Rpt_Yr_Lag > r$thresh)
    if (nrow(pool) == 0) return(tibble())
    ll <- sample(pool$Rpt_Yr_Lag, size = r$N_IBNR_b, replace = TRUE,
                 prob = pool$prop / sum(pool$prop))
    tibble(claim_id   = paste0("IBNR_", r$Acc_Yr, "_", seq_len(r$N_IBNR_b)),
           Acc_Yr     = as.integer(r$Acc_Yr),
           Rpt_Yr_Lag = as.integer(ll))
  })
  state_ibnr <- ibnr_synth_b %>%
    transmute(claim_num = claim_id, Acc_Yr, Rpt_Yr_Lag,
              dev = as.integer(Rpt_Yr_Lag - 1L),
              cum = 0, prob_open = 1.0)
  R_ibnr_ay <- if (nrow(state_ibnr) > 0) proj(state_ibnr)
               else tibble(Acc_Yr = ay_range, R = 0)
  
  R_total_ay <- R_rbns_ay %>%
    rename(R_RBNS = R) %>%
    left_join(R_ibnr_ay %>% rename(R_IBNR = R), by = "Acc_Yr") %>%
    mutate(R_IBNR = replace_na(R_IBNR, 0),
           R_TOT  = R_RBNS + R_IBNR)
  
  list(by_ay  = R_total_ay,
       totals = c(RBNS  = sum(R_total_ay$R_RBNS),
                  IBNR  = sum(R_total_ay$R_IBNR),
                  TOTAL = sum(R_total_ay$R_TOT)))
}

# Bootstrap loop
B <- 100
claim_ids <- unique(claims$claim_num)
boot_results  <- matrix(NA, nrow = B, ncol = 3,
                        dimnames = list(NULL, c("RBNS", "IBNR", "TOTAL")))
boot_rf_by_ay <- matrix(NA_real_, nrow = B, ncol = length(ay_range),
                        dimnames = list(NULL, as.character(ay_range)))
cat(sprintf("\nRunning %d RF bootstrap replicates (with by-AY tracking)...\n", B))
t0 <- Sys.time()
for (b in seq_len(B)) {
  set.seed(1000 + b)
  sampled_ids <- sample(claim_ids, length(claim_ids), replace = TRUE)
  boot_claims <- tibble(claim_num = sampled_ids,
                        boot_id = seq_along(sampled_ids)) %>%
    inner_join(claims, by = "claim_num", relationship = "many-to-many") %>%
    mutate(claim_num = paste0(claim_num, "_b", boot_id)) %>%
    select(-boot_id)
  out <- tryCatch(
    boot_one(boot_claims, seed = 2000 + b),
    error = function(e) { cat("  iter", b, "failed:", e$message, "\n"); NULL }
  )
  if (!is.null(out)) {
    boot_results[b, ]  <- out$totals
    boot_rf_by_ay[b, as.character(out$by_ay$Acc_Yr)] <- out$by_ay$R_TOT
  }
  cat(sprintf("  iter %2d/%d: TOTAL = $%s   elapsed: %.1f min\n",
              b, B,
              format(round(boot_results[b, "TOTAL"]), big.mark = ","),
              as.numeric(difftime(Sys.time(), t0, units = "mins"))))
}

# Save to disk to avoid re-running the bootstrap (it's slow)
saveRDS(boot_rf_by_ay, "boot_rf_by_ay.rds")
saveRDS(boot_results,  "boot_rf_totals.rds")
```

## Variance-based credibility combination of CL and RF

$$
\hat{R}_i^{blend} = (1-z_i)\hat{R}_i^{CL,adj} + z_i \hat{R}_i^{RF+IBNR}
$$

```{r combination CL-RF, echo=TRUE}
# Bootstrap variances by accident year
var_cl <- apply(boot_by_ay,    2, var, na.rm = TRUE)
var_rf <- apply(boot_rf_by_ay, 2, var, na.rm = TRUE)

# weights
z_var <- var_cl / (var_cl + var_rf)
z_tbl <- tibble(
  Acc_Yr = as.integer(names(var_cl)),
  var_CL = var_cl,
  var_RF = var_rf,
  z      = z_var
) %>%
  mutate(z = ifelse(is.na(z) | is.nan(z), 0.5, z))
cat("\n=== Variance-based bootstrap weights (Neuhaus-Mack) ===\n")
print(z_tbl %>% mutate(across(starts_with("var_"), ~ round(. / 1e12, 2)),
                       z = round(z, 3)))

# Blend
bf_blend_var <- z_tbl %>%
  left_join(reserve_infl       %>% select(Acc_Yr, R_CL_adj  = R_hat_infl),  by = "Acc_Yr") %>%
  left_join(reserves_rf_total  %>% select(Acc_Yr, R_RF_IBNR = R_RF_total), by = "Acc_Yr") %>%
  mutate(
    R_CL_adj   = replace_na(R_CL_adj,  0),
    R_RF_IBNR  = replace_na(R_RF_IBNR, 0),
    #       (1 - z) = CL weight,  z = RF weight
    R_blend_var = (1 - z) * R_CL_adj + z * R_RF_IBNR
  )
cat(sprintf("\n  Total CL inflation-adj: %s USD\n",
            format(round(sum(bf_blend_var$R_CL_adj)),    big.mark = ",")))
cat(sprintf("  Total RF + IBNR:        %s USD\n",
            format(round(sum(bf_blend_var$R_RF_IBNR)),   big.mark = ",")))
cat(sprintf("  Total VAR-based blend:  %s USD\n",
            format(round(sum(bf_blend_var$R_blend_var)), big.mark = ",")))
print(bf_blend_var %>%
        select(Acc_Yr, z, R_CL_adj, R_RF_IBNR, R_blend_var) %>%
        mutate(z = round(z, 3),
               across(starts_with("R_"), ~ round(. / 1e6, 2))))
```

```{r plot blend variance, echo=TRUE}
# Weights by AY
p_z_var <- bf_blend_var %>%
  mutate(`Weight on RF (z)`     = z,
         `Weight on CL (1-z)`   = 1 - z) %>%
  select(Acc_Yr, `Weight on CL (1-z)`, `Weight on RF (z)`) %>%
  pivot_longer(-Acc_Yr, names_to = "Component", values_to = "Weight") %>%
  ggplot(aes(factor(Acc_Yr), Weight, fill = Component)) +
  geom_col() +
  scale_fill_manual(values = c("Weight on CL (1-z)" = "#1b4f72",
                               "Weight on RF (z)"   = "#e67e22")) +
  labs(title    = "Variance-based credibility weights by accident year",
       subtitle = "z = Var_CL / (Var_CL + Var_RF)",
       x = "Accident Year", y = "Weight") +
  theme(legend.position = "bottom")
ggsave("figs/bf_weights_var.pdf", p_z_var, width = 7, height = 4)
print(p_z_var)

# Comparison of the three blend types (var-based vs 1/F vs components)
p_compare_blends <- bf_blend_var %>%
  select(Acc_Yr, R_CL_adj, R_RF_IBNR, R_blend_var) %>%
  pivot_longer(-Acc_Yr, names_to = "Method", values_to = "Reserve") %>%
  mutate(Method = recode(Method,
                         R_CL_adj    = "CL inflation-adj",
                         R_RF_IBNR   = "RF + IBNR",
                         R_blend_var = "Blend z = Var-based")) %>%
  ggplot(aes(factor(Acc_Yr), Reserve / 1e6, fill = Method)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("CL inflation-adj"      = "#1b4f72",
                               "RF + IBNR"             = "#e67e22",
                               "Blend z = Var-based"   = "#8e44ad")) +
  scale_y_continuous(labels = comma) +
  labs(title = "Reserve estimates: CL, RF, and variance-based blend",
       x = "Accident Year", y = "Reserve (USD millions)") +
  theme(legend.position = "bottom")
ggsave("figs/blends_compare.pdf", p_compare_blends, width = 8, height = 4.5)
print(p_compare_blends)
```

```{r blend var total summary, echo=TRUE}
#  Final summary: total reserve comparison across the three methods
R_tot_CL    <- sum(bf_blend_var$R_CL_adj)
R_tot_RF    <- sum(bf_blend_var$R_RF_IBNR)
R_tot_blend <- sum(bf_blend_var$R_blend_var)

cat("\n══════ TOTAL RESERVE — method comparison ═══════\n")
cat(sprintf("  CL inflation-adjusted    : %s USD  (%.0f M)\n",
            format(round(R_tot_CL),    big.mark = ","), R_tot_CL    / 1e6))
cat(sprintf("  RF + IBNR                : %s USD  (%.0f M)\n",
            format(round(R_tot_RF),    big.mark = ","), R_tot_RF    / 1e6))
cat(sprintf("  Blend z = Var-based      : %s USD  (%.0f M)\n",
            format(round(R_tot_blend), big.mark = ","), R_tot_blend / 1e6))
cat(sprintf("  ─────────────────────────────────────────────\n"))
cat(sprintf("  Blend position within the CL–RF range: %.0f%%\n",
            100 * (R_tot_blend - R_tot_CL) / (R_tot_RF - R_tot_CL)))
```

### other graphs

```{r other graphs, echo=TRUE}
fig_cl_vs_rf_rbns <- reserve_infl %>%
  select(Acc_Yr, `CL inflation-adjusted` = R_hat_infl) %>%
  full_join(
    reserves_ml %>% select(Acc_Yr, `RF (RBNS only, no IBNR)` = R_hat_RF),
    by = "Acc_Yr"
  ) %>%
  mutate(across(where(is.numeric), ~ replace_na(., 0))) %>%
  pivot_longer(-Acc_Yr, names_to = "Method", values_to = "Reserve") %>%
  ggplot(aes(factor(Acc_Yr), Reserve / 1e6, fill = Method)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c(
    "CL inflation-adjusted" = "#1b4f72",
    "RF (RBNS only, no IBNR)" = "#e67e22"
  )) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "CL inflation-adjusted vs RF (RBNS only) by accident year",
    x = "Accident Year",
    y = "Reserve (USD millions)"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
ggsave("figs/fig6_cl_vs_rf_rbns.pdf",
       fig_cl_vs_rf_rbns, width = 7, height = 4)
print(fig_cl_vs_rf_rbns)

reserve_summary <- data.frame(
  Method = c(
    "CL infl-adjusted",
    "RF RBNS",
    "RF + IBNR",
    "CL moderate",
    "CL full",
    "RF + Cox + IBNR",
    "Var-based blend"
  ),
  Reserve_M = c(
    sum(reserve_infl$R_hat_infl, na.rm = TRUE) / 1e6,
    sum(reserves_ml$R_hat_RF, na.rm = TRUE) / 1e6,
    sum(reserves_rf_total$R_RF_total, na.rm = TRUE) / 1e6,
    sum(reserves_scenarios$R_moderate, na.rm = TRUE) / 1e6,
    sum(reserves_scenarios$R_full, na.rm = TRUE) / 1e6,
    sum(reserves_cox_total$R_Cox_total, na.rm = TRUE) / 1e6,
    sum(bf_blend_var$R_blend_var, na.rm = TRUE) / 1e6
  )
)
reserve_summary <- reserve_summary[order(reserve_summary$Reserve_M), ]
reserve_summary$Method <- factor(reserve_summary$Method, levels = reserve_summary$Method)
fig10_convergence <- ggplot(reserve_summary,
                            aes(x = Method, y = Reserve_M, fill = Method)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = paste0(round(Reserve_M), "M")),
            vjust = -0.4, fontface = "bold", size = 4) +
  scale_y_continuous(
    labels = comma,
    limits = c(0, max(reserve_summary$Reserve_M) * 1.15)
  ) +
  scale_fill_manual(values = c(
    "CL infl-adjusted" = "#1b4f72",
    "RF RBNS" = "#e67e22",
    "RF + IBNR" = "#c0392b",
    "CL moderate" = "#8e44ad",
    "CL full" = "#d35400",
    "RF + Cox + IBNR" = "#27ae60",
    "Var-based blend" = "#16a085"
  )) +
  labs(
    title = "Reserve estimates across main methods",
    x = NULL,
    y = "Total reserve (USD millions)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 25, hjust = 1)
  )
ggsave("figs/fig10_convergence.png",
       fig10_convergence, width = 8, height = 5, dpi = 300)
print(fig10_convergence)
```

```{r, echo=TRUE}
reserve_summary <- data.frame(
  Method = c(
    "CL infl-adjusted",
    "RF RBNS",
    "RF + IBNR",
    "CL moderate",
    "CL full",
    "Var-based blend"
  ),
  Reserve_M = c(
    sum(reserve_infl$R_hat_infl, na.rm = TRUE) / 1e6,
    sum(reserves_ml$R_hat_RF, na.rm = TRUE) / 1e6,
    sum(reserves_rf_total$R_RF_total, na.rm = TRUE) / 1e6,
    sum(reserves_scenarios$R_moderate, na.rm = TRUE) / 1e6,
    sum(reserves_scenarios$R_full, na.rm = TRUE) / 1e6,
    sum(bf_blend_var$R_blend_var, na.rm = TRUE) / 1e6
  )
)
reserve_summary <- reserve_summary[order(reserve_summary$Reserve_M), ]
reserve_summary$Method <- factor(reserve_summary$Method, levels = reserve_summary$Method)
fig10_conv <- ggplot(reserve_summary,
                            aes(x = Method, y = Reserve_M, fill = Method)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = paste0(round(Reserve_M), "M")),
            vjust = -0.4, fontface = "bold", size = 4) +
  scale_y_continuous(
    labels = comma,
    limits = c(0, max(reserve_summary$Reserve_M) * 1.15)
  ) +
  scale_fill_manual(values = c(
    "CL infl-adjusted" = "#1b4f72",
    "RF RBNS" = "#e67e22",
    "RF + IBNR" = "#c0392b",
    "CL moderate" = "#8e44ad",
    "CL full" = "#d35400",
    "Var-based blend" = "#16a085"
  )) +
  labs(
    title = "Reserve estimates across main methods",
    x = NULL,
    y = "Total reserve (USD millions)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 25, hjust = 1)
  )
print(fig10_conv)
```

